### 2022-01-04 tikz.tex
new test

    \documentclass[tikz,border=0cm]{standalone}
    \usepackage{graphicx}
    \usepackage{tikz}
    \usetikzlibrary{positioning}
    \usetikzlibrary{arrows,snakes,shapes,fadings}
    \usepackage{fontspec}
    \setmainfont{Roboto}[
      Extension = .otf,
      UprightFont = *-Regular,
      BoldFont = *-Bold,
      ItalicFont = *-Italic,
      BoldItalicFont = *-BoldItalic,
    ]    \usepackage{etoolbox}

    \newcommand{\figwidth}{16cm} 
    \newcommand{\figheight}{10cm} 
    \newcommand{\fontselect}{\large\bf}

    \tikzset{
      labelNode/.style={anchor=south east, above left=0cm and 0cm of anchor, font=\fontselect}, 
      redAnchorNode/.style={anchor=center, left=-0.1cm of anchor, circle, fill=red, minimum size=0.2cm}, 
      graphicNode/.style={anchor=north west, below right=0cm and 0cm of anchor}, 
    }

    \newtoggle{draft}
        \toggletrue{draft}
        % \togglefalse{draft}

    \begin{document}
    \begin{tikzpicture}[every node/.style={inner sep=0pt}]

    \iftoggle{draft}{\node[anchor=south west, rectangle, draw, red, line width=2pt, minimum width=\figwidth, minimum height=\figheight] at (0,0) {};}
    {\node[anchor=south west, rectangle, fill=white, minimum width=\figwidth, minimum height=\figheight] at (0,0) {};};

    \node (anchor) at (0.5,9.35) {}; 
\node[labelNode] {A};
\node[graphicNode] {\includegraphics[scale=0.1]{content/graphics/approach/approach.pdf}

};
\iftoggle{draft}{\node [redAnchorNode] {};};
    \end{tikzpicture}
    \end{document}