### 2022-01-08 content/graphics/approach/logo.tex
Some work

\documentclass[tikz,border=0cm]{standalone}
\usepackage{graphicx}
\usepackage{tikz}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows,snakes,shapes,fadings,calc}
\usepackage{fontspec}
\setmainfont{Roboto}[
  Extension = .otf,
  UprightFont = *-Regular,
  BoldFont = *-Bold,
  ItalicFont = *-Italic,
  BoldItalicFont = *-BoldItalic,
]    
\usepackage{etoolbox}

\newcommand{\figwidth}{7cm} 
\newcommand{\figheight}{10cm} 
\newcommand{\fontselect}{\large\bf}

\begin{document}
\begin{tikzpicture}

  % Include the image in a node
  \node [
      above right,
      inner sep=0] (image) at (0,0) {\includegraphics[width=\textwidth]{approach.png}};

  % Create scope with normalized axes
  \node [above right,inner sep=0,text width=\textwidth ,align=center] at (0,13) {\Huge \textbf{BILACE}};
  \node [above right,inner sep=0,text width=\textwidth ,align=center] at (0,12.5) {\normalsize \textbf{B}eam-\textbf{I}mageshift for \textbf{L}arge \textbf{A}rea \textbf{C}ryo-\textbf{E}lectron microscopy}; 
  \end{tikzpicture}
\end{document}