### 2022-03-03 figures/initmatching.tex

    \documentclass[tikz,border=0cm]{standalone}
    \usepackage{graphicx}
    \usepackage{tikz}
    \usepackage{contour}
    \contourlength{0.6pt}
    \usetikzlibrary{positioning}
    \usetikzlibrary{arrows,snakes,shapes,fadings}
    \usetikzlibrary{fit}
    \usetikzlibrary{backgrounds}
    \usepackage{fontspec}
    \setmainfont{Roboto}[
      Extension = .otf,
      UprightFont = *-Regular,
      BoldFont = *-Bold,
      ItalicFont = *-Italic,
      BoldItalicFont = *-BoldItalic,
    ]    \usepackage{etoolbox} 
    \usepackage{enumitem}
    \setlist{nosep,leftmargin=*}
    \newcommand{\figwidth}{19cm} 
    \newcommand{\figheight}{10cm} 
    \newcommand{\fontselect}{\large\bf} 
    \newcommand{\fontannot}{\bf}

    \tikzset{
      labelNode/.style={anchor=south east, above left=0cm and 0cm of anchor, font=\fontselect}, 
      annotNode/.style={anchor=center, font=\fontannot},
      redAnchorNode/.style={anchor=center, left=-0.1cm of anchor, circle, fill=red, minimum size=0.2cm}, 
      graphicNode/.style={anchor=north west, below right=0cm and 0cm of anchor}, 
      titleNode/.style={anchor=south west, above right=0cm and 0cm of anchor}, 
      mnNode/.style={below left=0.1cm and 0cm of rect.south east, font=\fontannot},
      block/.style = {rectangle, draw, fill=gray!20, text width=12em, text centered, rounded corners, minimum height=2em,inner sep=0.5em},
      line/.style = {draw, -latex'}
    }

    \newtoggle{draft}
     \togglefalse{draft}
     %\togglefalse{draft}

    \begin{document}
    \begin{tikzpicture}[
      every node/.style={inner sep=0pt},
      box/.style = {draw=#1, line width=0.8mm,inner sep=0.0mm}
    ]

    \iftoggle{draft}{\node[anchor=south west, rectangle, draw, red, line width=2pt, minimum width=\figwidth, minimum height=\figheight] at (0,0) {};}
    {
      \begin{scope}[on background layer]

      \node[anchor=south west, rectangle, fill=white, minimum width=\figwidth, minimum height=\figheight] at (0,0) {};
      \end{scope}
    };

    \node (anchor) at (0.5,9.35) {}; 
\node[labelNode] {A};
\node[graphicNode] {\includegraphics[height=9cm]{content/images/the_lamella.png}};

\node[annotNode] at (3.5,8.35) {\contour{white}{Nuc}};
\node[annotNode] at (1.75,4.0) {\contour{white}{Nuc}};

\node[annotNode] at (3.2,4.7) {\contour{white}{M}};
\node[annotNode] at (2.75,2.25) {\contour{white}{M}};

\node[annotNode] at (1.5,8.0) {\contour{white}{GrCyt}};
\node[annotNode] at (4.5,3.5) {\contour{white}{GrCyt}};

\node (rect) at (1.1,7.5) [anchor=center,draw,thick,minimum width=0.85cm,minimum height=0.6cm,color=blue!60] {};
\node[mnNode, color=blue!60] {\contour{white}{2}};

\node (rect) at (3.2,7.3) [anchor=center,draw,thick,minimum width=0.85cm,minimum height=0.6cm,color=blue!60] {};
\node[mnNode, color=blue!60] {\contour{white}{0}};

\node (rect) at (4.5,6.25) [anchor=center,draw,thick,minimum width=0.85cm,minimum height=0.6cm,color=red] {};
\node[mnNode, color=red] {\contour{white}{506}};

\node (rect) at (3.4,5.5) [anchor=center,draw,thick,minimum width=0.85cm,minimum height=0.6cm,color=blue!60] {};
\node[mnNode, color=blue!60] {\contour{white}{16}};

\node (rect) at (2.75,3.5) [anchor=center,draw,thick,minimum width=0.85cm,minimum height=0.6cm,color=green] {};
\node[mnNode, color=green] {\contour{white}{30}};

\node (rect) at (2.65,2.35) [anchor=center,draw,thick,minimum width=0.85cm,minimum height=0.6cm,color=orange] {};
\node[mnNode, color=orange] {\contour{white}{17}};

\node (rect) at (1.25,1.5) [anchor=center,draw,thick,minimum width=0.85cm,minimum height=0.6cm,color=blue!60] {};
\node[mnNode, color=blue!60] {\contour{white}{8}};

\node (rect) at (4.0,2.2) [anchor=center,draw,thick,minimum width=0.85cm,minimum height=0.6cm,color=magenta] {};
\node[mnNode, color=magenta] {\contour{white}{52}};

\node (anchor) at (6.5,9.35) {}; 
\node[labelNode] {B};
\node (img) [graphicNode]  {\includegraphics[height=4cm]{content/images/30_matches.png}};
\node (n3) [box=green, fit=(img)] {};

\node (anchor) at (12.75,9.35) {}; 
\node[labelNode] {C};
\node (img) [graphicNode]  {\includegraphics[height=4cm]{content/images/506_matches.png}};
\node (n3) [box=red, fit=(img)] {};

\node (anchor) at (6.5,4.75) {}; 
\node[labelNode] {D};
\node (img) [graphicNode]  {\includegraphics[height=4cm]{content/images/17_matches.png}};
\node (n3) [box=orange, fit=(img)] {};

\node (anchor) at (12.75,4.75) {}; 
\node[labelNode] {E};
\node (img) [graphicNode]  {\includegraphics[height=4cm]{content/images/52_matches.png}};
\node (n3) [box=magenta, fit=(img)] {};

%%\iftoggle{draft}{\node [redAnchorNode] {};};

    \iftoggle{draft}{\draw[lightgray,step=1] (0,0) grid (\figwidth,\figheight);}{};
    \end{tikzpicture}
    \end{document}