### 2022-04-07 figures/approach.tex
Minor figure [ci skip]
This build is based on
https://github.com/jojoelfe/deco_lace_template_matching_manuscript/commit/56ce938e8d7fb4ca49b0ca90d4ceca7386ae3b0f.

This commit was created by the following CI build and job:
https://github.com/jojoelfe/deco_lace_template_matching_manuscript/commit/56ce938e8d7fb4ca49b0ca90d4ceca7386ae3b0f/checks
https://github.com/jojoelfe/deco_lace_template_matching_manuscript/actions/runs/2110566960

    \documentclass[tikz,border=0cm]{standalone}
    \usepackage{graphicx}
    \usepackage{tikz}
    \usepackage{contour}
    \contourlength{0.6pt}
    \usetikzlibrary{positioning}
    \usetikzlibrary{arrows,snakes,shapes,fadings}
    \usetikzlibrary{fit}
    \usetikzlibrary{backgrounds}
    \usepackage{fontspec}
    \setmainfont{Roboto}[
      Extension = .otf,
      UprightFont = *-Regular,
      BoldFont = *-Bold,
      ItalicFont = *-Italic,
      BoldItalicFont = *-BoldItalic,
    ]    \usepackage{etoolbox} 
    \usepackage{enumitem}
    \setlist{nosep,leftmargin=*}
    \newcommand{\figwidth}{20cm} 
    \newcommand{\figheight}{20cm} 
    \newcommand{\fontselect}{\large\bf} 
    \newcommand{\fontannot}{\bf}

    \tikzset{
      labelNode/.style={anchor=south east, above left=0cm and 0cm of anchor, font=\fontselect}, 
      annotNode/.style={anchor=center, font=\fontannot},
      redAnchorNode/.style={anchor=center, left=-0.1cm of anchor, circle, fill=red, minimum size=0.2cm}, 
      graphicNode/.style={anchor=north west, below right=0cm and 0cm of anchor}, 
      mnNode/.style={below left=0.1cm and 0cm of rect.south east, font=\fontannot},
      block/.style = {rectangle, draw, fill=gray!20, text width=12em, text centered, rounded corners, minimum height=2em,inner sep=0.5em},
      line/.style = {draw, -latex'}
    }

    \newtoggle{draft}
     \togglefalse{draft}
     %\togglefalse{draft}

    \begin{document}
    \begin{tikzpicture}[
      every node/.style={inner sep=0pt},
      box/.style = {draw=#1, line width=0.8mm,inner sep=0.0mm}
    ]

    \iftoggle{draft}{\node[anchor=south west, rectangle, draw, red, line width=2pt, minimum width=\figwidth, minimum height=\figheight] at (0,0) {};}
    {
      \begin{scope}[on background layer]

      \node[anchor=south west, rectangle, fill=white, minimum width=\figwidth, minimum height=\figheight] at (0,0) {};
      \end{scope}
    };

    \node (anchor) at (0.5,19.35) {}; 
\node[labelNode] {A};
\node[graphicNode] {\includegraphics[width=5cm]{content/graphics/approach/approach.png}};
\draw [densely dotted, red,thick] (3,19.5) -- (3,15.0);
\draw [densely dotted, red,thick] (3,14.2) -- (3,11.4);

\tikzset{dimen/.style={<->,>=latex,thin,every rectangle node/.style={above right=0.2cm and -0.3cm,fill=white,midway,font=\sffamily}}}

\draw (3.55,15.4) -- ++(0,1.5) coordinate (D1) -- +(0,5pt);
\draw (4.15,15.4) -- ++(0,1.5) coordinate (D2) -- +(0,5pt);
\draw [dimen] (D1) -- (D2) node {$D_{Beam}$};

\draw [->, very thick] (3,15) -- ++(0.5,-0.2)  coordinate (S1) node[midway,below left] {$IS_X$} ;
\draw [->, very thick] (S1) -- (3.85,15.4) node[midway,below right] {$IS_Y$};

\node (anchor) at (6.0,19.35) {}; 
\node[labelNode] {B};

% Place nodes
\node [block] (init) at (8.0,18.35) {\textbf{START} \\ Collect overview image};
\node [block, below = 0.3cm of init.south] (identify) {Setup hexagonal grid covering lamella area};
\node [block, below = 0.3cm of identify.south] (mutate) {Beam-imageshift to tile n=i};
\node [block, below = 0.3cm of mutate.south] (transfer) {Capture exposure and determine defocus by CTF fitting};
\node [block, below = 0.3cm of transfer.south] (evaluate) {Adjust defocus by difference between expected and measured value};

% Draw edges
\path [line] (init) -- (identify);
\path [line] (identify) -- (mutate);
\path [line] (mutate) -- (transfer);
\path [line] (transfer) -- (evaluate);
\path [line] (evaluate) -|  ([xshift=0.5cm, yshift=0cm]transfer.east) |- (mutate);

\node (anchor) at (11.5,19.35) {}; 
\node[labelNode] {C};
\node[graphicNode] {\includegraphics[width=8cm]{content/images/ac_strategy.png}};
\iftoggle{draft}{\node [redAnchorNode] {};};

\node (anchor) at (0.5,9.5) {}; 
\node[labelNode] {D};
\node[graphicNode] (euc) {\includegraphics[height=4.0cm]{content/images/fringebeam.png}};
\node[above = 0cm and 0cm of euc.north] {Eucentric Focus};

\iftoggle{draft}{\node [redAnchorNode] {};};

\node (anchor) at (5.5,9.5) {}; 
\node[labelNode] {E};
\node[graphicNode] (fff) {\includegraphics[height=4.0cm]{content/images/nofringebeam.png}};
\node[above = 0cm and 0cm of fff.north] {Fringe-free Focus};
\iftoggle{draft}{\node [redAnchorNode] {};};

\node (anchor) at (10.5,9.5) {}; 
\node[labelNode] {F};
\node[anchor=north west, below right=-0.5cm and 0cm of anchor] {\includegraphics[height=4.8cm]{content/code/img/defocusplot.pdf}};
\iftoggle{draft}{\node [redAnchorNode] {};};

\node (anchor) at (0.5,4.5) {}; 
\node[labelNode] {G};
\node[graphicNode] {\includegraphics[height=4.5cm]{content/images/euc_lamella1_view_morph.png}};

\node (anchor) at (5,4.5) {}; 
\node[labelNode] {H};
\node[graphicNode] {\includegraphics[height=4.5cm]{content/images/euc_lamella1_1000px.png}};

\node (anchor) at (9.5,4.5) {}; 
\node[labelNode] {I};
\node[graphicNode] {\includegraphics[height=4.5cm]{content/images/fff_lamella4_view_morphed.png}};

\node (anchor) at (14,4.5) {}; 
\node[labelNode] {J};
\node[graphicNode] {\includegraphics[height=4.5cm]{content/images/fff_lamella4_1000px.png}};

    \iftoggle{draft}{\draw[lightgray,step=1] (0,0) grid (\figwidth,\figheight);}{};
    \end{tikzpicture}
    \end{document}