### 2022-04-07 figures/crop_unblur.tex

    \documentclass[tikz,border=0cm]{standalone}
    \usepackage{graphicx}
    \usepackage{tikz}
    \usepackage{contour}
    \contourlength{0.6pt}
    \usetikzlibrary{positioning}
    \usetikzlibrary{arrows,snakes,shapes,fadings}
    \usetikzlibrary{fit}
    \usetikzlibrary{backgrounds}
    \usepackage{fontspec}
    \setmainfont{Roboto}[
      Extension = .otf,
      UprightFont = *-Regular,
      BoldFont = *-Bold,
      ItalicFont = *-Italic,
      BoldItalicFont = *-BoldItalic,
    ]    \usepackage{etoolbox} 
    \usepackage{enumitem}
    \setlist{nosep,leftmargin=*}
    \newcommand{\figwidth}{20.0cm} 
    \newcommand{\figheight}{10cm} 
    \newcommand{\fontselect}{\large\bf} 
    \newcommand{\fontannot}{\bf}

    \tikzset{
      labelNode/.style={anchor=south east, above left=0cm and 0cm of anchor, font=\fontselect}, 
      annotNode/.style={anchor=center, font=\fontannot},
      redAnchorNode/.style={anchor=center, left=-0.1cm of anchor, circle, fill=red, minimum size=0.2cm}, 
      graphicNode/.style={anchor=north west, below right=0cm and 0cm of anchor}, 
      mnNode/.style={below left=0.1cm and 0cm of rect.south east, font=\fontannot},
      block/.style = {rectangle, draw, fill=gray!20, text width=12em, text centered, rounded corners, minimum height=2em,inner sep=0.5em},
      line/.style = {draw, -latex'}
    }

    \newtoggle{draft}
     \togglefalse{draft}
     %\togglefalse{draft}

    \begin{document}
    \begin{tikzpicture}[
      every node/.style={inner sep=0pt},
      box/.style = {draw=#1, line width=0.8mm,inner sep=0.0mm}
    ]

    \iftoggle{draft}{\node[anchor=south west, rectangle, draw, red, line width=2pt, minimum width=\figwidth, minimum height=\figheight] at (0,0) {};}
    {
      \begin{scope}[on background layer]

      \node[anchor=south west, rectangle, fill=white, minimum width=\figwidth, minimum height=\figheight] at (0,0) {};
      \end{scope}
    };

    \def\panelheight{4.6cm}

\node (anchor) at (0.5,9.5) {}; 
\node[labelNode] {A};
\node[graphicNode] (panel) {\includegraphics[width=5cm]{content/images/crop_unblur_initial.png}};
\node (box1) [box=red, fit=(panel), dashed] {};

\node (anchor) at ([shift={(0.0,-0.5cm)}]panel.south west) {}; 
\node[graphicNode] (panel) {\includegraphics[width=5cm]{content/code/img/crop_unblur_motiontrace_initial.png}};

\node (anchor) at ([shift={(0.0,-0.5cm)}]panel.south west) {}; 
\node[graphicNode] (panel) {\includegraphics[width=5cm]{content/code/img/crop_unblur_mip_initial_alignment.png}};

\node (anchor) at ([shift={(1.5,0.0cm)}]0.5,9.5 -| panel.north east) {}; 
\node[labelNode] {B};
\node[graphicNode] (panel) {\includegraphics[width=5cm]{content/images/crop_unblur_cropped.png}};
\node (box2) [box=red, fit=(panel), dashed] {};

\node (anchor) at ([shift={(0.0,-0.5cm)}]panel.south west) {}; 
\node[graphicNode] (panel) {\includegraphics[width=5cm]{content/code/img/crop_unblur_motiontrace_cropped.png}};

\node (anchor) at ([shift={(0.0,-0.5cm)}]panel.south west) {}; 
\node[graphicNode] (panel) {\includegraphics[width=5cm]{content/code/img/crop_unblur_mip_cropped_alignment.png}};

\node (anchor) at ([shift={(1.5,0.0cm)}]0.5,9.5 -| panel.north east) {}; 
\node[labelNode] {C};
\node[graphicNode] (panel) {\includegraphics[width=5cm]{content/images/crop_unblur_initial.png}};
\draw[red, dashed, line width=0.8mm] ([shift={(-0.7,-1.2cm)}]panel.north east) -- 
                                     ([shift={(0.7,-1.2cm)}]panel.north west) --
                                     ([shift={(0.7,1.2cm)}]panel.south west) --
                                     ([shift={(-0.7,1.2cm)}]panel.south east) --
                                     ([shift={(-0.7,-1.2cm)}]panel.north east);

\node (anchor) at ([shift={(0.0,-0.5cm)}]panel.south west) {}; 
\node[graphicNode] (panel) {\includegraphics[width=5cm]{content/code/img/crop_unblur_motiontrace_final.png}};

\node (anchor) at ([shift={(0.0,-0.5cm)}]panel.south west) {}; 
\node[graphicNode] (panel) {\includegraphics[width=5cm]{content/code/img/crop_unblur_mip_final_alignment.png}};

    \iftoggle{draft}{\draw[lightgray,step=1] (0,0) grid (\figwidth,\figheight);}{};
    \end{tikzpicture}
    \end{document}