### 2022-05-19 figures/assembly.tex

    \documentclass[tikz,border=0cm]{standalone}
    \usepackage{graphicx}
    \usepackage{tikz}
    \usepackage{contour}
    \contourlength{0.6pt}
    \usetikzlibrary{positioning}
    \usetikzlibrary{arrows,snakes,shapes,fadings}
    \usetikzlibrary{fit}
    \usetikzlibrary{calc}
    \def\centerarc[#1](#2)(#3:#4:#5){ \draw[#1] ($(#2)+({#5*cos(#3)},{#5*sin(#3)})$) arc (#3:#4:#5); }
    \usetikzlibrary{3d}
    \usetikzlibrary{backgrounds}
    \usepackage{fontspec}
    \setmainfont{Roboto}[
      Extension = .otf,
      UprightFont = *-Regular,
      BoldFont = *-Bold,
      ItalicFont = *-Italic,
      BoldItalicFont = *-BoldItalic,
    ]    \usepackage{etoolbox} 
    \usepackage{enumitem}
    \setlist{nosep,leftmargin=*}
    \newcommand{\figwidth}{15cm} 
    \newcommand{\figheight}{5cm} 
    \newcommand{\fontselect}{\large\bf} 
    \newcommand{\fontannot}{\bf}

    \tikzset{
      labelNode/.style={anchor=south east, above left=0cm and 0cm of anchor, font=\fontselect}, 
      annotNode/.style={anchor=center, font=\fontannot},
      redAnchorNode/.style={anchor=center, left=-0.1cm of anchor, circle, fill=red, minimum size=0.2cm}, 
      graphicNode/.style={anchor=north west, below right=0cm and 0cm of anchor}, 
      mnNode/.style={below left=0.1cm and 0cm of rect.south east, font=\fontannot},
      block/.style = {rectangle, draw, fill=gray!20, text width=12em, text centered, rounded corners, minimum height=2em,inner sep=0.5em},
      line/.style = {draw, -latex'}
    }

    \newtoggle{draft}
     \togglefalse{draft}
     %\togglefalse{draft}

    \begin{document}
    \begin{tikzpicture}[
      every node/.style={inner sep=0pt},
      box/.style = {draw=#1, line width=0.8mm,inner sep=0.0mm}
    ]

    \iftoggle{draft}{\node[anchor=south west, rectangle, draw, red, line width=2pt, minimum width=\figwidth, minimum height=\figheight] at (0,0) {};}
    {
      \begin{scope}[on background layer]

      \node[anchor=south west, rectangle, fill=white, minimum width=\figwidth, minimum height=\figheight] at (0,0) {};
      \end{scope}
    };

    \def\panelheight{6.0cm}

\node (anchor) at (0.5,13.0) {}; 
\node[labelNode] {A};
\node[graphicNode] (img) {\includegraphics[height=\panelheight,trim=0.0 0 0cm 0, clip]{content/images/euc_lamella1_view.png}};
% Title
\node[anchor=center,align=center] at ([shift={(0.0cm,0.6cm)}]img.north) {\baselineskip=8pt   \\ 38Å pixel size \\ 300μm defocus};
% scalebar - physical height is 12810 nm
\draw [line width=1.55mm,white] ([xshift=0.5cm, yshift=0.3cm]img.south west) -- ([xshift=0.5cm + \panelheight / 12.81 , yshift=0.3cm]img.south west);
\draw [line width=0.7mm,black] ([xshift=0.5cm, yshift=0.3cm]img.south west) -- ([xshift=0.5cm + \panelheight / 12.81, yshift=0.3cm]img.south west);

\node (anchor) at (5.5,13.0) {}; 
\node[labelNode] {B};
\node[graphicNode] (img) {\includegraphics[height=6.0cm,trim=0 0 0cm 0, clip]{content/images/euc_lamella1_assm.png}};
\node[anchor=center,align=center] at ([shift={(0.0cm,0.7cm)}]img.north) {\baselineskip=8pt Eucentric Focus \\ 1.76Å pixel size \\ 800nm defocus};
% Red Box
\node (rect) at (6.52,10.78) [anchor=center,draw,thick,minimum width=0.375cm,minimum height=0.469cm,color=red] {};
% scalebar - physical height is 12810 nm
\draw [line width=1.55mm,white] ([xshift=0.5cm, yshift=0.3cm]img.south west) -- ([xshift=0.5cm + \panelheight / 12.81 , yshift=0.3cm]img.south west);
\draw [line width=0.7mm,black] ([xshift=0.5cm, yshift=0.3cm]img.south west) -- ([xshift=0.5cm + \panelheight / 12.81, yshift=0.3cm]img.south west);

\node (anchor) at (10.6,13.0) {}; 
\node[labelNode] {C};
% height is 750 nm
\node[graphicNode] (image) {\includegraphics[height=6.0cm]{content/images/euc_lam1_zoomin.png}};
\node (n3) [box=red, fit=(image)] {};
\draw [line width=1.55mm,white] ([xshift=0.5cm, yshift=0.3cm]image.south west) -- ([xshift=1.3cm, yshift=0.3cm]image.south west);
\draw [line width=0.7mm,black] ([xshift=0.5cm, yshift=0.3cm]image.south west) -- ([xshift=1.3cm, yshift=0.3cm]image.south west);

\node (anchor) at (0.5,5.0) {}; 
\node[labelNode] {D};
\node[graphicNode] (img) {\includegraphics[height=6.0cm,trim=0 0 0cm 0, clip]{content/images/fff_lamella4_view.png}};
\node[anchor=center,align=center] at ([shift={(0.0cm,0.6cm)}]img.north) {\baselineskip=8pt  \\ 38Å pixel size \\ 300μm defocus};
% scalebar - physical height is 9337.5 nm
\draw [line width=1.55mm,white] ([xshift=0.5cm, yshift=0.3cm]img.south west) -- ([xshift=0.5cm + \panelheight / 9.3375 , yshift=0.3cm]img.south west);
\draw [line width=0.7mm,black] ([xshift=0.5cm, yshift=0.3cm]img.south west) -- ([xshift=0.5cm + \panelheight / 9.3375, yshift=0.3cm]img.south west);

\node (anchor) at (5.8,5.0) {}; 
\node[labelNode] {E};
\node[graphicNode] (img) {\includegraphics[height=6.0cm,trim=0 0 0cm 0, clip]{content/images/fff_lamella4_assm.png}};
\node[anchor=center,align=center] at ([shift={(0.0cm,0.7cm)}]img.north) {\baselineskip=2pt Fringe-free Focus \\ 1.76Å pixel size \\ 800nm defocus};
\node (rect) at (8.438,3.83) [anchor=center,draw,thick,minimum width=0.375cm,minimum height=0.469cm,color=red] {};
% scalebar - physical height is 9337.5 nm
\draw [line width=1.55mm,white] ([xshift=0.5cm, yshift=0.3cm]img.south west) -- ([xshift=0.5cm + \panelheight / 9.3375 , yshift=0.3cm]img.south west);
\draw [line width=0.7mm,black] ([xshift=0.5cm, yshift=0.3cm]img.south west) -- ([xshift=0.5cm + \panelheight / 9.3375, yshift=0.3cm]img.south west);

\node (anchor) at (11.0,5.0) {}; 
\node[labelNode] {F};
% height is 750 nm
\node[graphicNode] (image) {\includegraphics[height=6.0cm]{content/images/fff_lam4_zoomin.png}};
\node (n3) [box=red, fit=(image)] {};
\draw [line width=1.55mm,white] ([xshift=0.5cm, yshift=0.3cm]image.south west) -- ([xshift=1.3cm, yshift=0.3cm]image.south west);
\draw [line width=0.7mm,black] ([xshift=0.5cm, yshift=0.3cm]image.south west) -- ([xshift=1.3cm, yshift=0.3cm]image.south west);
    \iftoggle{draft}{\draw[lightgray,step=1] (0,0) grid (\figwidth,\figheight);}{};
    \end{tikzpicture}
    \end{document}