### 2022-05-20 figures/lamella_images.tex

    \documentclass[tikz,border=0cm]{standalone}
    \usepackage{graphicx}
    \usepackage{tikz}
    \usepackage{contour}
    \contourlength{0.6pt}
    \usetikzlibrary{positioning}
    \usetikzlibrary{arrows,snakes,shapes,fadings}
    \usetikzlibrary{fit}
    \usetikzlibrary{calc}
    \def\centerarc[#1](#2)(#3:#4:#5){ \draw[#1] ($(#2)+({#5*cos(#3)},{#5*sin(#3)})$) arc (#3:#4:#5); }
    \usetikzlibrary{3d}
    \usetikzlibrary{backgrounds}
    \usepackage{fontspec}
    \setmainfont{Roboto}[
      Extension = .otf,
      UprightFont = *-Regular,
      BoldFont = *-Bold,
      ItalicFont = *-Italic,
      BoldItalicFont = *-BoldItalic,
    ]    \usepackage{etoolbox} 
    \usepackage{enumitem}
    \setlist{nosep,leftmargin=*}
    \newcommand{\figwidth}{20.0cm} 
    \newcommand{\figheight}{5cm} 
    \newcommand{\fontselect}{\large\bf} 
    \newcommand{\fontannot}{\bf}

    \tikzset{
      labelNode/.style={anchor=south east, above left=0cm and 0cm of anchor, font=\fontselect}, 
      annotNode/.style={anchor=center, font=\fontannot},
      redAnchorNode/.style={anchor=center, left=-0.1cm of anchor, circle, fill=red, minimum size=0.2cm}, 
      graphicNode/.style={anchor=north west, below right=0cm and 0cm of anchor}, 
      mnNode/.style={below left=0.1cm and 0cm of rect.south east, font=\fontannot},
      block/.style = {rectangle, draw, fill=gray!20, text width=12em, text centered, rounded corners, minimum height=2em,inner sep=0.5em},
      line/.style = {draw, -latex'}
    }

    \newtoggle{draft}
     \togglefalse{draft}
     %\togglefalse{draft}

    \begin{document}
    \begin{tikzpicture}[
      every node/.style={inner sep=0pt},
      box/.style = {draw=#1, line width=0.8mm,inner sep=0.0mm}
    ]

    \iftoggle{draft}{\node[anchor=south west, rectangle, draw, red, line width=2pt, minimum width=\figwidth, minimum height=\figheight] at (0,0) {};}
    {
      \begin{scope}[on background layer]

      \node[anchor=south west, rectangle, fill=white, minimum width=\figwidth, minimum height=\figheight] at (0,0) {};
      \end{scope}
    };

    \newlength{\panelheight}
\setlength{\panelheight}{4.4cm}

\node (anchor) at (0.5,4.5) {}; 
\node[labelNode] {A};
\node[graphicNode] (panel) {\includegraphics[height=\panelheight]{content/images/euc_lamella1_view.png}};
% Title
\node[anchor=center,align=center] at ([shift={(0.0cm,0.3cm)}]panel.north) {Lamella$_\textrm{EUC}$ 1};
% scalebar - physical height is 12810 nm
\draw [line width=1.55mm,white] ([xshift=0.5cm, yshift=0.3cm]panel.south west) -- ([xshift=0.5cm + \panelheight / 12.81 , yshift=0.3cm]panel.south west);
\draw [line width=0.7mm,black] ([xshift=0.5cm, yshift=0.3cm]panel.south west) -- ([xshift=0.5cm + \panelheight / 12.81, yshift=0.3cm]panel.south west);
\node (anchor) at ([shift={(0.5cm,0.0cm)}]panel.north east) {}; 
\node[graphicNode] (panel) {\includegraphics[height=\panelheight]{content/images/euc_lamella2_view.png}};
% Title
\node[anchor=center,align=center] at ([shift={(0.0cm,0.3cm)}]panel.north) {Lamella$_\textrm{EUC}$ 2};
% scalebar - physical height is 7027.5 nm
\draw [line width=1.55mm,white] ([xshift=0.5cm, yshift=0.3cm]panel.south west) -- ([xshift=0.5cm + \panelheight / 7.0275 , yshift=0.3cm]panel.south west);
\draw [line width=0.7mm,black] ([xshift=0.5cm, yshift=0.3cm]panel.south west) -- ([xshift=0.5cm + \panelheight / 7.0275, yshift=0.3cm]panel.south west);
\node (anchor) at ([shift={(0.5cm,0.0cm)}]panel.north east) {}; 
\node[graphicNode] (panel) {\includegraphics[height=\panelheight]{content/images/euc_lamella3_view.png}};
% Title
\node[anchor=center,align=center] at ([shift={(0.0cm,0.3cm)}]panel.north) {Lamella$_\textrm{EUC}$ 3};
% scalebar - physical height is 9615 nm
\draw [line width=1.55mm,white] ([xshift=0.5cm, yshift=0.3cm]panel.south west) -- ([xshift=0.5cm + \panelheight / 9.615 , yshift=0.3cm]panel.south west);
\draw [line width=0.7mm,black] ([xshift=0.5cm, yshift=0.3cm]panel.south west) -- ([xshift=0.5cm + \panelheight / 9.615, yshift=0.3cm]panel.south west);
\node (anchor) at ([shift={(0.5cm,0.0cm)}]panel.north east) {}; 
\node[graphicNode] (panel) {\includegraphics[height=\panelheight]{content/images/euc_lamella4_view.png}};
% Title
\node[anchor=center,align=center] at ([shift={(0.0cm,0.3cm)}]panel.north) {Lamella$_\textrm{EUC}$ 4};
% scalebar - physical height is 8625 nm
\draw [line width=1.55mm,white] ([xshift=0.5cm, yshift=0.3cm]panel.south west) -- ([xshift=0.5cm + \panelheight / 8.625 , yshift=0.3cm]panel.south west);
\draw [line width=0.7mm,black] ([xshift=0.5cm, yshift=0.3cm]panel.south west) -- ([xshift=0.5cm + \panelheight / 8.625, yshift=0.3cm]panel.south west);

\node (anchor) at ([shift={(0.0cm,-0.5cm)}]0.5,0 |- panel.south) {}; 
\node[graphicNode]  (panel) {\includegraphics[height=\panelheight]{content/images/fff_lamella1_view.png}};
% Title
\node[anchor=center,align=center] at ([shift={(0.0cm,0.3cm)}]panel.north) {Lamella$_\textrm{FFF}$ 1};
% scalebar - physical height is 8032.5 nm
\draw [line width=1.55mm,white] ([xshift=0.5cm, yshift=0.3cm]panel.south west) -- ([xshift=0.5cm + \panelheight / 8.0325 , yshift=0.3cm]panel.south west);
\draw [line width=0.7mm,black] ([xshift=0.5cm, yshift=0.3cm]panel.south west) -- ([xshift=0.5cm + \panelheight / 8.0325, yshift=0.3cm]panel.south west);
\node (anchor) at ([shift={(0.5cm,0.0cm)}]panel.north east) {}; 
\node[graphicNode]  (panel) {\includegraphics[height=\panelheight]{content/images/fff_lamella2_view.png}};
% Title
\node[anchor=center,align=center] at ([shift={(0.0cm,0.3cm)}]panel.north) {Lamella$_\textrm{FFF}$ 2};
% scalebar - physical height is 10000.5 nm
\draw [line width=1.55mm,white] ([xshift=0.5cm, yshift=0.3cm]panel.south west) -- ([xshift=0.5cm + \panelheight / 10.0005 , yshift=0.3cm]panel.south west);
\draw [line width=0.7mm,black] ([xshift=0.5cm, yshift=0.3cm]panel.south west) -- ([xshift=0.5cm + \panelheight / 10.0005, yshift=0.3cm]panel.south west);
\node (anchor) at ([shift={(0.5cm,0.0cm)}]panel.north east) {}; 
\node[graphicNode]  (panel) {\includegraphics[height=\panelheight]{content/images/fff_lamella3_view.png}};
% Title
\node[anchor=center,align=center] at ([shift={(0.0cm,0.3cm)}]panel.north) {Lamella$_\textrm{FFF}$ 3};
% scalebar - physical height is 9975 nm
\draw [line width=1.55mm,white] ([xshift=0.5cm, yshift=0.3cm]panel.south west) -- ([xshift=0.5cm + \panelheight / 9.975 , yshift=0.3cm]panel.south west);
\draw [line width=0.7mm,black] ([xshift=0.5cm, yshift=0.3cm]panel.south west) -- ([xshift=0.5cm + \panelheight / 9.975, yshift=0.3cm]panel.south west);
\node (anchor) at ([shift={(0.5cm,0.0cm)}]panel.north east) {}; 
\node[graphicNode]  (panel) {\includegraphics[height=\panelheight]{content/images/fff_lamella4_view.png}};
% Title
\node[anchor=center,align=center] at ([shift={(0.0cm,0.3cm)}]panel.north) {Lamella$_\textrm{FFF}$ 4};
% scalebar - physical height is 9337.5 nm
\draw [line width=1.55mm,white] ([xshift=0.5cm, yshift=0.3cm]panel.south west) -- ([xshift=0.5cm + \panelheight / 9.3375 , yshift=0.3cm]panel.south west);
\draw [line width=0.7mm,black] ([xshift=0.5cm, yshift=0.3cm]panel.south west) -- ([xshift=0.5cm + \panelheight / 9.3375, yshift=0.3cm]panel.south west);

\node (anchor) at ([shift={(0.0cm,-1.0cm)}]0.5,0 |- panel.south) {}; 
\node[labelNode] {B};
\node[graphicNode] (panel) {\includegraphics[height=\panelheight]{content/images/euc_lamella1_assm.png}};
% Title
\node[anchor=center,align=center] at ([shift={(0.0cm,0.3cm)}]panel.north) {Lamella$_\textrm{EUC}$ 1: 909 tiles};
% scalebar - physical height is 12810 nm
\draw [line width=1.55mm,white] ([xshift=0.5cm, yshift=0.3cm]panel.south west) -- ([xshift=0.5cm + \panelheight / 12.81 , yshift=0.3cm]panel.south west);
\draw [line width=0.7mm,black] ([xshift=0.5cm, yshift=0.3cm]panel.south west) -- ([xshift=0.5cm + \panelheight / 12.81, yshift=0.3cm]panel.south west);
\node (anchor) at ([shift={(0.5cm,0.0cm)}]panel.north east) {}; 
\node[graphicNode] (panel) {\includegraphics[height=\panelheight]{content/images/euc_lamella2_assm.png}};
% Title
\node[anchor=center,align=center] at ([shift={(0.0cm,0.3cm)}]panel.north) {Lamella$_\textrm{EUC}$ 2: 451 tiles};
% scalebar - physical height is 7027.5 nm
\draw [line width=1.55mm,white] ([xshift=0.5cm, yshift=0.3cm]panel.south west) -- ([xshift=0.5cm + \panelheight / 7.0275 , yshift=0.3cm]panel.south west);
\draw [line width=0.7mm,black] ([xshift=0.5cm, yshift=0.3cm]panel.south west) -- ([xshift=0.5cm + \panelheight / 7.0275, yshift=0.3cm]panel.south west);
\node (anchor) at ([shift={(0.5cm,0.0cm)}]panel.north east) {}; 
\node[graphicNode] (panel) {\includegraphics[height=\panelheight]{content/images/euc_lamella3_assm.png}};
% Title
\node[anchor=center,align=center] at ([shift={(0.0cm,0.3cm)}]panel.north) {Lamella$_\textrm{EUC}$ 3: 513 tiles};
% scalebar - physical height is 9615 nm
\draw [line width=1.55mm,white] ([xshift=0.5cm, yshift=0.3cm]panel.south west) -- ([xshift=0.5cm + \panelheight / 9.615 , yshift=0.3cm]panel.south west);
\draw [line width=0.7mm,black] ([xshift=0.5cm, yshift=0.3cm]panel.south west) -- ([xshift=0.5cm + \panelheight / 9.615, yshift=0.3cm]panel.south west);
\node (anchor) at ([shift={(0.5cm,0.0cm)}]panel.north east) {}; 
\node[graphicNode] (panel) {\includegraphics[height=\panelheight]{content/images/euc_lamella4_assm.png}};
% Title
\node[anchor=center,align=center] at ([shift={(0.0cm,0.3cm)}]panel.north) {Lamella$_\textrm{EUC}$ 4: 497 tiles};
% scalebar - physical height is 8625 nm
\draw [line width=1.55mm,white] ([xshift=0.5cm, yshift=0.3cm]panel.south west) -- ([xshift=0.5cm + \panelheight / 8.625 , yshift=0.3cm]panel.south west);
\draw [line width=0.7mm,black] ([xshift=0.5cm, yshift=0.3cm]panel.south west) -- ([xshift=0.5cm + \panelheight / 8.625, yshift=0.3cm]panel.south west);

\node (anchor) at ([shift={(0.0cm,-0.5cm)}]0.5,0 |- panel.south) {}; 
\node[graphicNode]  (panel) {\includegraphics[height=\panelheight]{content/images/fff_lamella1_assm.png}};
% Title
\node[anchor=center,align=center] at ([shift={(0.0cm,0.3cm)}]panel.north) {Lamella$_\textrm{FFF}$ 1: 438 tiles};
% scalebar - physical height is 10000.5 nm
\draw [line width=1.55mm,white] ([xshift=0.5cm, yshift=0.3cm]panel.south west) -- ([xshift=0.5cm + \panelheight / 10.0005 , yshift=0.3cm]panel.south west);
\draw [line width=0.7mm,black] ([xshift=0.5cm, yshift=0.3cm]panel.south west) -- ([xshift=0.5cm + \panelheight / 10.0005, yshift=0.3cm]panel.south west);
\node (anchor) at ([shift={(0.5cm,0.0cm)}]panel.north east) {}; 
\node[graphicNode]  (panel) {\includegraphics[height=\panelheight]{content/images/fff_lamella2_assm.png}};
% Title
\node[anchor=center,align=center] at ([shift={(0.0cm,0.3cm)}]panel.north) {Lamella$_\textrm{FFF}$ 2: 474 tiles};
% scalebar - physical height is 10000.5 nm
\draw [line width=1.55mm,white] ([xshift=0.5cm, yshift=0.3cm]panel.south west) -- ([xshift=0.5cm + \panelheight / 10.0005 , yshift=0.3cm]panel.south west);
\draw [line width=0.7mm,black] ([xshift=0.5cm, yshift=0.3cm]panel.south west) -- ([xshift=0.5cm + \panelheight / 10.0005, yshift=0.3cm]panel.south west);
\node (anchor) at ([shift={(0.5cm,0.0cm)}]panel.north east) {}; 
\node[graphicNode]  (panel) {\includegraphics[height=\panelheight]{content/images/fff_lamella3_assm.png}};
% Title
\node[anchor=center,align=center] at ([shift={(0.0cm,0.3cm)}]panel.north) {Lamella$_\textrm{FFF}$ 3: 501 tiles};
% scalebar - physical height is 9975 nm
\draw [line width=1.55mm,white] ([xshift=0.5cm, yshift=0.3cm]panel.south west) -- ([xshift=0.5cm + \panelheight / 9.975 , yshift=0.3cm]panel.south west);
\draw [line width=0.7mm,black] ([xshift=0.5cm, yshift=0.3cm]panel.south west) -- ([xshift=0.5cm + \panelheight / 9.975, yshift=0.3cm]panel.south west);
\node (anchor) at ([shift={(0.5cm,0.0cm)}]panel.north east) {}; 
\node[graphicNode]  (panel) {\includegraphics[height=\panelheight]{content/images/fff_lamella4_assm.png}};
% Title
\node[anchor=center,align=center] at ([shift={(0.0cm,0.3cm)}]panel.north) {Lamella$_\textrm{FFF}$ 4: 425 tiles};
% scalebar - physical height is 9337.5 nm
\draw [line width=1.55mm,white] ([xshift=0.5cm, yshift=0.3cm]panel.south west) -- ([xshift=0.5cm + \panelheight / 9.3375 , yshift=0.3cm]panel.south west);
\draw [line width=0.7mm,black] ([xshift=0.5cm, yshift=0.3cm]panel.south west) -- ([xshift=0.5cm + \panelheight / 9.3375, yshift=0.3cm]panel.south west);

\newlength{\padding}
\setlength{\padding}{0.2cm}
\node (anchor) at ([shift={(0.0cm,-1.0cm)}]0.5,0 |- panel.south) {}; 
\node[labelNode] {C};
\node[graphicNode] (panel) {\includegraphics[height=0.75\panelheight]{content/images/granules/zap004.png}};
\node (anchor) at ([shift={(\padding,0.0cm)}]panel.north east) {}; 
\node[graphicNode] (panel) {\includegraphics[height=0.75\panelheight]{content/images/granules/zap012.png}};
\node (anchor) at ([shift={(\padding,0.0cm)}]panel.north east) {}; 
\node[graphicNode] (panel) {\includegraphics[height=0.75\panelheight]{content/images/granules/zap015.png}};
\node (anchor) at ([shift={(\padding,0.0cm)}]panel.north east) {}; 
\node[graphicNode] (panel) {\includegraphics[height=0.75\panelheight]{content/images/granules/zap020.png}};
\node (anchor) at ([shift={(\padding,0.0cm)}]panel.north east) {}; 
\node[graphicNode] (panel) {\includegraphics[height=0.75\panelheight]{content/images/granules/zap006.png}};

\node (anchor) at ([shift={(0.0cm,-\padding)}]0.5,0 |- panel.south) {}; 
\node[graphicNode]  (panel) {\includegraphics[height=0.75\panelheight]{content/images/granules/zap021.png}};
\node (anchor) at ([shift={(\padding,0.0cm)}]panel.north east) {}; 
\node[graphicNode]  (panel) {\includegraphics[height=0.75\panelheight]{content/images/granules/zap022.png}};
\node (anchor) at ([shift={(\padding,0.0cm)}]panel.north east) {}; 
\node[graphicNode]  (panel) {\includegraphics[height=0.75\panelheight]{content/images/granules/zap027.png}};
\node (anchor) at ([shift={(\padding,0.0cm)}]panel.north east) {}; 
\node[graphicNode]  (panel) {\includegraphics[height=0.75\panelheight]{content/images/granules/zap029.png}};
\node (anchor) at ([shift={(\padding,0.0cm)}]panel.north east) {}; 
\node[graphicNode]  (panel) {\includegraphics[height=0.75\panelheight]{content/images/granules/zap023.png}};
% scalebar - physical height is 375 nm
\draw [line width=1.55mm,white] ([xshift=0.5cm, yshift=0.3cm]panel.south west) -- ([xshift=0.5cm + 0.75\panelheight / 3.75 , yshift=0.3cm]panel.south west);
\draw [line width=0.7mm,black] ([xshift=0.5cm, yshift=0.3cm]panel.south west) -- ([xshift=0.5cm + 0.75\panelheight / 3.75, yshift=0.3cm]panel.south west);

    \iftoggle{draft}{\draw[lightgray,step=1] (0,0) grid (\figwidth,\figheight);}{};
    \end{tikzpicture}
    \end{document}