### 2022-05-20 figures/matching_euc.tex

    \documentclass[tikz,border=0cm]{standalone}
    \usepackage{graphicx}
    \usepackage{tikz}
    \usepackage{contour}
    \contourlength{0.6pt}
    \usetikzlibrary{positioning}
    \usetikzlibrary{arrows,snakes,shapes,fadings}
    \usetikzlibrary{fit}
    \usetikzlibrary{calc}
    \def\centerarc[#1](#2)(#3:#4:#5){ \draw[#1] ($(#2)+({#5*cos(#3)},{#5*sin(#3)})$) arc (#3:#4:#5); }
    \usetikzlibrary{3d}
    \usetikzlibrary{backgrounds}
    \usepackage{fontspec}
    \setmainfont{Roboto}[
      Extension = .otf,
      UprightFont = *-Regular,
      BoldFont = *-Bold,
      ItalicFont = *-Italic,
      BoldItalicFont = *-BoldItalic,
    ]    \usepackage{etoolbox} 
    \usepackage{enumitem}
    \setlist{nosep,leftmargin=*}
    \newcommand{\figwidth}{17cm} 
    \newcommand{\figheight}{10cm} 
    \newcommand{\fontselect}{\large\bf} 
    \newcommand{\fontannot}{\bf}

    \tikzset{
      labelNode/.style={anchor=south east, above left=0cm and 0cm of anchor, font=\fontselect}, 
      annotNode/.style={anchor=center, font=\fontannot},
      redAnchorNode/.style={anchor=center, left=-0.1cm of anchor, circle, fill=red, minimum size=0.2cm}, 
      graphicNode/.style={anchor=north west, below right=0cm and 0cm of anchor}, 
      mnNode/.style={below left=0.1cm and 0cm of rect.south east, font=\fontannot},
      block/.style = {rectangle, draw, fill=gray!20, text width=12em, text centered, rounded corners, minimum height=2em,inner sep=0.5em},
      line/.style = {draw, -latex'}
    }

    \newtoggle{draft}
     \togglefalse{draft}
     %\togglefalse{draft}

    \begin{document}
    \begin{tikzpicture}[
      every node/.style={inner sep=0pt},
      box/.style = {draw=#1, line width=0.8mm,inner sep=0.0mm}
    ]

    \iftoggle{draft}{\node[anchor=south west, rectangle, draw, red, line width=2pt, minimum width=\figwidth, minimum height=\figheight] at (0,0) {};}
    {
      \begin{scope}[on background layer]

      \node[anchor=south west, rectangle, fill=white, minimum width=\figwidth, minimum height=\figheight] at (0,0) {};
      \end{scope}
    };

    \newlength{\panelheight}
\setlength{\panelheight}{12cm}

\node (anchor) at (0.5,9.5) {}; 
\node[labelNode] {A};
\node[graphicNode] (assm) {\includegraphics[height=\panelheight]{content/images/euc_lamella1_assm.png}};
\node[graphicNode] {\includegraphics[height=\panelheight]{content/images/euc_lamella1.png}};
\node (rect) at (7.4,3.5) [anchor=center,draw,thick,minimum width=0.3cm,minimum height=0.3cm,color=red] {};
\node (rect) at (1.75,4.5) [anchor=center,draw,thick,minimum width=0.3cm,minimum height=0.3cm,color=green] {};
\node (rect) at (7.2,2.15) [anchor=center,draw,thick,minimum width=0.3cm,minimum height=0.3cm,color=blue!60] {};
\node (rect) at (6.45,6.0) [anchor=center,draw,thick,minimum width=0.47cm,minimum height=0.47cm,color=cyan] {};
% scalebar - physical height is 12810 nm
\draw [line width=1.55mm,white] ([xshift=0.5cm, yshift=0.3cm]assm.south west) -- ([xshift=0.5cm + \panelheight / 12.81 , yshift=0.3cm]assm.south west);
\draw [line width=0.7mm,black] ([xshift=0.5cm, yshift=0.3cm]assm.south west) -- ([xshift=0.5cm + \panelheight / 12.81, yshift=0.3cm]assm.south west);

\node (anchor) at (11.0,9.5) {}; 
\node[labelNode] {B};
\node[anchor=north west, below right=0cm and 0.0cm of anchor] (side) {\includegraphics[height=12cm, trim=20.5cm 0 2cm 0]{content/images/euc_lamella1_side.png}};

\draw [->, thick] ($ (assm.east) + (0.1,0) $) -- node[above=0.2cm, align=center] (c) {} ($ (side.west) + (0.2,0) $);

\centerarc [canvas is zx plane at y=0.75, ->](c.center)(-90:180:0.15);
\draw [very thick] ($ (c.center) + (0,-0.1) $) -- ($ (c.center) + (0,0.2) $);
\node at ($ (c.center) + (0.5,0) $) {90°};

\newlength{\panelheightm}
\setlength{\panelheightm}{2.8cm}
\node (anchor) at (13.0,9.5) {}; 
\node[labelNode] {C};
\node[graphicNode] (pan) {\includegraphics[height=\panelheightm]{content/images/euc_lamella1_detail1.png}};
\node (n3) [box=red, fit=(pan)] {};
% scalebar - physical height is 300 nm
\draw [line width=1.55mm,white] ([xshift=0.2cm, yshift=0.2cm]pan.south west) -- ([xshift=0.2cm + \panelheightm / 3 , yshift=0.2cm]pan.south west);
\draw [line width=0.7mm,black] ([xshift=0.2cm, yshift=0.2cm]pan.south west) -- ([xshift=0.2cm + \panelheightm / 3, yshift=0.2cm]pan.south west);

\node (anchor) at (13.0,6.5) {}; 
\node[labelNode] {D};
\node[graphicNode] (pan) {\includegraphics[height=\panelheightm]{content/images/euc_lamella1_detail2.png}};
\node (n3) [box=green, fit=(pan)] {};

% scalebar - physical height is 300 nm
\draw [line width=1.55mm,white] ([xshift=0.2cm, yshift=0.2cm]pan.south west) -- ([xshift=0.2cm + \panelheightm / 3 , yshift=0.2cm]pan.south west);
\draw [line width=0.7mm,black] ([xshift=0.2cm, yshift=0.2cm]pan.south west) -- ([xshift=0.2cm + \panelheightm / 3, yshift=0.2cm]pan.south west);

\node (anchor) at (13.,3.5) {}; 
\node[labelNode] {E};
\node[graphicNode] (pan) {\includegraphics[height=\panelheightm]{content/images/euc_lamella1_detail3.png}};
\node (n3) [box=blue!60, fit=(pan)] {};

% scalebar - physical height is 300 nm
\draw [line width=1.55mm,white] ([xshift=0.2cm, yshift=0.2cm]pan.south west) -- ([xshift=0.2cm + \panelheightm / 3 , yshift=0.2cm]pan.south west);
\draw [line width=0.7mm,black] ([xshift=0.2cm, yshift=0.2cm]pan.south west) -- ([xshift=0.2cm + \panelheightm / 3, yshift=0.2cm]pan.south west);

\node (anchor) at (13.0,0.5) {}; 
\node[labelNode] {F};
\node[graphicNode] (pan) {\includegraphics[height=\panelheightm]{content/images/euc_lamella1_detail4.png}};
\node (n3) [box=cyan, fit=(pan)] {};

% scalebar - physical height is 500 nm
\draw [line width=1.55mm,white] ([xshift=0.2cm, yshift=0.2cm]pan.south west) -- ([xshift=0.2cm + \panelheightm / 5 , yshift=0.2cm]pan.south west);
\draw [line width=0.7mm,black] ([xshift=0.2cm, yshift=0.2cm]pan.south west) -- ([xshift=0.2cm + \panelheightm / 5, yshift=0.2cm]pan.south west);

    \iftoggle{draft}{\draw[lightgray,step=1] (0,0) grid (\figwidth,\figheight);}{};
    \end{tikzpicture}
    \end{document}