### 2022-11-14 elife_submission/figures/deco_lace_workflow.tex
Minor adjustments to submission

    \documentclass[tikz,border=0cm]{standalone}
    \usepackage{graphicx}
    \usepackage{tikz}
    \usepackage{contour}
    \contourlength{0.6pt}
    \usetikzlibrary{positioning}
    \usetikzlibrary{arrows,snakes,shapes,fadings}
    \usetikzlibrary{fit}
    \usetikzlibrary{calc}
    \def\centerarc[#1](#2)(#3:#4:#5){ \draw[#1] ($(#2)+({#5*cos(#3)},{#5*sin(#3)})$) arc (#3:#4:#5); }
    \usetikzlibrary{3d}
    \usetikzlibrary{backgrounds}
    \usepackage{fontspec}
    \setmainfont{Roboto}[
      Extension = .otf,
      UprightFont = *-Regular,
      BoldFont = *-Bold,
      ItalicFont = *-Italic,
      BoldItalicFont = *-BoldItalic,
    ]    \usepackage{etoolbox} 
    \usepackage{enumitem}
    \setlist{nosep,leftmargin=*}
    \newcommand{\figwidth}{18cm} 
    \newcommand{\figheight}{18cm} 
    \newcommand{\fontselect}{\large\bf} 
    \newcommand{\fontannot}{\bf}

    \tikzset{
      labelNode/.style={anchor=south east, above left=0cm and 0cm of anchor, font=\fontselect}, 
      annotNode/.style={anchor=center, font=\fontannot},
      redAnchorNode/.style={anchor=center, left=-0.1cm of anchor, circle, fill=red, minimum size=0.2cm}, 
      graphicNode/.style={anchor=north west, below right=0cm and 0cm of anchor}, 
      mnNode/.style={below left=0.1cm and 0cm of rect.south east, font=\fontannot},
      block/.style = {rectangle, draw, fill=gray!20, text width=12em, text centered, rounded corners, minimum height=2em,inner sep=0.5em},
      line/.style = {draw, -latex'}
    }

    \newtoggle{draft}
     \togglefalse{draft}
     %\togglefalse{draft}

    \begin{document}
    \begin{tikzpicture}[
      every node/.style={inner sep=0pt},
      box/.style = {draw=#1, line width=0.8mm,inner sep=0.0mm}
    ]

    \iftoggle{draft}{\node[anchor=south west, rectangle, draw, red, line width=2pt, minimum width=\figwidth, minimum height=\figheight] at (0,0) {};}
    {
      \begin{scope}[on background layer]

      \node[anchor=south west, rectangle, fill=white, minimum width=\figwidth, minimum height=\figheight] at (0,0) {};
      \end{scope}
    };

    \begin{scope}[
          blocks/.style = {rectangle, draw, fill=blue!20, text width=15em, align=center, rounded corners, minimum height=2em,inner sep=0.5em},
          inputs/.style = {rectangle,draw,fill=cyan!20, inner sep=0.5em},
          every path/.style={line width=1pt}
]
\node [inputs,anchor=north] (struc) at (3.0,17.5) {\textbf{6SWA Structure}};

\node [inputs,anchor=north] (movies) at (9.0,17.5) {\textbf{Movies}};
\node [inputs,anchor=north] (meta) at (15.0,17.5) {\textbf{SerialEM Metadata}};

\node [inputs,anchor=north] (over) at (3.0,1.5) {\textbf{Overview Image}};

\node [blocks,anchor=north] (simu) at (3.0,15.5) {\textbf{Generate Template} \\ \textit{simulate}
\begin{itemize}
    \item   Generate density map at 1.5{\AA}  pixel-size
\end{itemize}};

\node [blocks,anchor=north] (mc) at (9.0,15.5) {\textbf{Motion correction} \\ \textit{unblur\_decolace}
\begin{itemize}
    \item   Motion correction using central area of movie
    \item   Mask out unilluminated areas and replace with noise
\end{itemize}};

\node [inputs,below = 0.5 cm of mc.south] (images) {\textbf{Images}};

\node [blocks,below = 0.5 cm of images.south] (ctf) {\textbf{Ctf estimation} \\ \textit{ctffind4} 
\begin{itemize}
    \item Standard CTF estimation
\end{itemize}};
\node [blocks,below = 0.7 cm of ctf.south] (match) {\textbf{Template matching} \\ \textit{match\_template} \\ Template matching using 6swa as template};

\begin{scope}[on background layer]

\node[draw,very thick, dashed, inner sep=0.5em, rounded corners, fill=blue!10,
    fit=(mc) (ctf) (match)] (gui) {};
\end{scope}
\node[below left=0.1cm and 0cm of gui.south east, font=\fontannot] {cisTEM GUI};

\node [blocks,below = 1.4 cm of match.south] (refine) {\textbf{Refine matches} \\ \textit{refine\_matches.py/refine\_template} \\ Refine matches and compile list of all matches in a lamella};

\node [inputs,below = 0.5 cm of refine.south] (refstar) {\textbf{matches\_in\_tiles.star}};

\node [blocks,anchor=center] (assem) at (images.east -| meta.south) {\textbf{Assemble montage} \\ \textit{assemble\_montage.py} 
\begin{itemize}
    \item Create list of all tiles and their respective coordinates
    \item Create binned montage image
\end{itemize}};

\node [inputs,below = 0.5 cm of assem.south] (monstar) {\textbf{montage.star}};

\node [blocks,anchor=center] (assemma) at (refstar.east -| monstar.south) {\textbf{Assemble matches} \\ \textit{assemble\_matches.py} 
\begin{itemize}
    \item Transfer coordinates of matches into the montage coordinate system
\end{itemize}};

\node [inputs,below = 0.5 cm of assemma.south] (monmastar) {\textbf{matches\_in\_montage.star}};

%% Draw lines
\draw[->] (struc) -- (simu);

\draw[->] (movies) -- (mc);
\draw[->] (mc) -- (images);
\draw[->] (images) -- (ctf);
\draw[->] (ctf) -- (match);
\draw[->] (match) -- (refine);

\draw[->] (meta) -- (assem);
\draw[->] (images) -- (assem);

\draw[->] (simu) |- (match);
\draw[->] (simu) |- (refine);

\draw[->] (assem) -- (monstar);
\draw[->] (monstar) -- (assemma);

\draw[->] (refine) -- (refstar);
\draw[->] (refstar) -- (assemma);
\draw[->] (assemma) -- (monmastar);

\end{scope}
    \iftoggle{draft}{\draw[lightgray,step=1] (0,0) grid (\figwidth,\figheight);}{};
    \end{tikzpicture}
    \end{document}