### 2022-11-14 elife_submission/manuscript.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Defocus Corrected Large Area Cryo-EM (DeCo-LACE) for Label-Free Detection of Molecules across Entire Cell Sections},
  pdfauthor={Johannes ElferichGiulia SchiroliDavid ScaddenNikolaus Grigorieff},
  pdflang={en-US},
  pdfkeywords={cryo-EM, visual protemics, ribosome},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
\usepackage[bidi=basic]{babel}
\else
\usepackage[bidi=default]{babel}
\fi
\babelprovide[main,import]{american}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
\usepackage[margin=2cm]{geometry}
\usepackage{setspace}
\doublespacing
\usepackage{lineno}
\linenumbers
\usepackage{cleveref}
\usepackage{fontspec}
\setmainfont{texgyrepagella}[
  Extension = .otf,
  UprightFont = *-regular,
  BoldFont = *-bold,
  ItalicFont = *-italic,
  BoldItalicFont = *-bolditalic,
]
\usepackage{float}
\floatplacement{figure}{H}
\usepackage{caption}
\captionsetup[figure]{font={stretch=1.2}}

%% pandoc-fignos: environment for tagged figures
\newenvironment{fignos:tagged-figure}[1][]{
    \let\oldthefigure\thefigure
    \let\oldtheHfigure\theHfigure
    \renewcommand{\thefigure}{#1}
    \renewcommand{\theHfigure}{#1}
  }{
    \let\thefigure\oldthefigure
    \let\theHfigure\oldtheHfigure
    \addtocounter{figure}{-1}
  }

%% pandoc-tablenos: required package
\usepackage{caption}

%% pandoc-tablenos: environment to disable table caption prefixes
\makeatletter
\newcounter{tableno}
\newenvironment{tablenos:no-prefix-table-caption}{
    \let\oldthetable\thetable
    \let\oldtheHtable\theHtable
    \renewcommand{\thetable}{tableno:\thetableno}
    \renewcommand{\theHtable}{tableno:\thetableno}
    \stepcounter{tableno}
    \captionsetup{labelformat=empty}

}{
    \captionsetup{labelformat=default}
    \let\thetable\oldthetable
    \let\theHtable\oldtheHtable
    \addtocounter{table}{-1}

}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\title{Defocus Corrected Large Area Cryo-EM (DeCo-LACE) for Label-Free Detection of Molecules across Entire Cell Sections}
\author{}
\date{}

\begin{document}
\maketitle

\hypertarget{authors}{%
\subsection{Authors}\label{authors}}

\begin{itemize}
\item
  \textbf{Johannes Elferich}
  \includegraphics[width=0.16667in,height=0.16667in]{images/envelope.pdf}\includegraphics[width=0.16667in,height=0.16667in]{images/orcid.pdf}
  \href{https://orcid.org/0000-0002-9911-706X}{0000-0002-9911-706X}

  RNA Therapeutics Institute, University of
  Massachusetts Chan Medical School, Worcester, MA, USA; Howard Hughes
  Medical Institute, USA
\item
  \textbf{Giulia Schiroli} \includegraphics[width=0.16667in,height=0.16667in]{images/orcid.pdf}
  \href{https://orcid.org/0000-0002-9812-8728}{0000-0002-9812-8728}

  Center for Regenerative Medicine, Massachusetts
  General Hospital, Boston, MA, USA; Harvard Stem Cell Institute,
  Cambridge, MA, USA; Department of Stem Cell and Regenerative
  Biology, Harvard University, Cambridge, MA, USA
\item
  \textbf{David Scadden} \includegraphics[width=0.16667in,height=0.16667in]{images/orcid.pdf}
  \href{https://orcid.org/0000-0001-9821-7133}{0000-0001-9821-7133}

  Center for Regenerative Medicine, Massachusetts
  General Hospital, Boston, MA, USA; Harvard Stem Cell Institute,
  Cambridge, MA, USA; Department of Stem Cell and Regenerative
  Biology, Harvard University, Cambridge, MA, USA
\item
  \textbf{Nikolaus Grigorieff}
  \includegraphics[width=0.16667in,height=0.16667in]{images/envelope.pdf}\includegraphics[width=0.16667in,height=0.16667in]{images/orcid.pdf}
  \href{https://orcid.org/0000-0002-1506-909X}{0000-0002-1506-909X}

  RNA Therapeutics Institute, University of
  Massachusetts Chan Medical School, Worcester, MA, USA; Howard Hughes
  Medical Institute, USA
\end{itemize}

\leavevmode\vadjust pre{\hypertarget{correspondence}{}}%
\includegraphics[width=0.16667in,height=0.16667in]{images/envelope.pdf} --- correspondence preferred via \href{https://github.com/jojoelfe/deco_lace_template_matching_manuscript/issues}{GitHub
Issues}.
Otherwise, address correspondence to \href{mailto:johannes.elferich@umassmed.edu}{\nolinkurl{johannes.elferich@umassmed.edu}}
and \href{mailto:niko@grigorieff.org}{\nolinkurl{niko@grigorieff.org}}.

\hypertarget{abstract}{%
\subsection{Abstract}\label{abstract}}

A major goal of biological imaging is localization of biomolecules
inside a cell. Fluorescence microscopy can localize biomolecules inside
whole cells and tissues, but its ability to count biomolecules and
accuracy of the spatial coordinates is limited by the wavelength of
visible light. Cryo-electron microscopy (cryo-EM) provides highly
accurate position and orientation information of biomolecules but is
often confined to small fields of view inside a cell, limiting
biological context. In this study we use a new data-acquisition scheme
called ``Defocus-Corrected Large-Area cryo-EM'' (DeCo-LACE) to collect
high-resolution images of entire sections (100 -- 250 nm thick lamellae)
of neutrophil-like mouse cells, representing 1-2\% of the total cellular
volume. We use 2D template matching (2DTM) to determine localization and
orientation of the large ribosomal subunit in these sections. These data
provide ``maps'' of ribosomes across entire sections of mammalian cells.
This high-throughput cryo-EM data collection approach together with 2DTM
will advance visual proteomics and provide biological insight that
cannot be obtained by other methods.

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

A major goal in understanding cellular processes is the knowledge of the
amounts, location, interactions, and conformations of biomolecules
inside the cell. This knowledge can be obtained by approaches broadly
divided into label- and label-free techniques. In label-dependent
techniques a probe is physically attached to a molecule of interest that
is able to be detected by its strong signal, such as a fluorescent
molecule. In label-free techniques, the physical properties of molecules
themselves are used for detection. An example for this is proteomics
using mass-spectrometry {[}\protect\hyperlink{ref-tSXIKPl7}{1}{]}. The advantage of
label-free techniques is that they can provide information over
thousands of molecules, while label-dependent techniques offer highly
specific information for a few molecules. in particular, spatial
information is primarily achieved using label-dependent techniques, such
as fluorescence microscopy {[}\protect\hyperlink{ref-VBmW7Aot}{2}{]}.

Cryo-electron microscopy (cryo-EM) has the potential to directly
visualize the arrangement of atoms that compose biomolecules inside of
cells, thereby allowing label-free detection with high spatial accuracy.
This has been called ``visual proteomics'' {[}\protect\hyperlink{ref-tGQ6TSUo}{3}{]}. Since
cryo-EM requires thin samples (\textless500nm), imaging of biomolecules inside
cells is restricted to small organisms, thin regions of cells, or
samples that have been suitably thinned. Thinning can be achieved either
by mechanical sectioning {[}\protect\hyperlink{ref-g8QavfwP}{4}{]} or by
milling using a focused ion beam (FIB) {[}\protect\hyperlink{ref-16IhS1Nc4}{5}{]}.
This complex workflow leads to a low throughput of cryo-EM imaging of
cells and is further limited by the fact that at the required
magnifications, typical fields of view (FOV) are very small compared to
mammalian cells, and the FOV achieved by label-dependent techniques such
as fluorescence light microscopy. The predominant cryo-EM technique for
the localization of biomolecules of defined size and shape inside cells
is cryo-electron tomography {[}\protect\hyperlink{ref-Rksh2dxu}{6}{]}. However,
the requirement of a tilt series at every imaged location and subsequent
image alignment, severely limits the throughput for molecular
localization.

An alternative approach is to identify molecules by their structural
``fingerprint'' in single projection using ``2D template-matching'' (2DTM)
{[}\protect\hyperlink{ref-Ynb3IP6I}{7},\protect\hyperlink{ref-18KGpXYPE}{8},\protect\hyperlink{ref-10bXZuF3G}{9}{]}.
In this method, a 3D model of a biomolecule is used as a template to
find 2D projections that match the molecules visible in the electron
micrographs. This method requires a projection search on a fine angular
grid, and the projections are used to find local cross-correlation peaks
with the micrograph. Since the location of a biomolecule in the
z-direction causes predictable aberrations to the projection image, this
method can be used to calculate complete 3D coordinates and orientations
of a biomolecule in a cellular sample {[}\protect\hyperlink{ref-18KGpXYPE}{8}{]}.

Here we apply 2DTM of the ribosome large subunit (LSU) to a
conditionally immortalized \emph{mus musculus} (mouse) cell line that gives
rise to functional mature neutrophils {[}\protect\hyperlink{ref-1B9Vt9eYu}{10}{]}.
We chose these cells because genetic defects in the ribosome machinery
often leads to hematopoietic disease {[}\protect\hyperlink{ref-gRoY21jY}{11}{]} and
direct quantification of ribosome location, number and conformational
states in hematopoietic cells could lead to new insight into
hematopoietic disease {[}\protect\hyperlink{ref-KAJ7221k}{12}{]}. To increase the
amount of collected data and to provide unbiased sampling of the whole
lamella, we devised a new data-acquisition scheme, ``Defocus-Corrected
Large Area Cryo-Electron microscopy'' (DeCo-LACE). 2DTM allows us to test
whether aberrations caused by large beam-image shifts and highly
condensed beams deteriorate the high-resolution signal. We find that
these aberrations do not impede LSU detection by 2DTM. The resulting
data provide a description of ribosome distribution in an entire
lamella, which represent 1-2\% of the cellular volume. We find a highly
heterogeneous density of ribosomes within the cell. Analysis of the
throughput in this method suggests that for the foreseeable future
computation will be the bottleneck for visual proteomics.

\hypertarget{results}{%
\subsection{Results}\label{results}}

\hypertarget{dtm-detects-large-ribosomal-subunits-in-cryo-fib-lamellae-of-mammalian-cells}{%
\subsubsection{2DTM detects large ribosomal subunits in cryo-FIB lamellae of mammalian cells}\label{dtm-detects-large-ribosomal-subunits-in-cryo-fib-lamellae-of-mammalian-cells}}

FIB-milled \emph{Saccharomyces cerevisiae} (yeast) cells are sufficiently
well preserved to permit localization of 60S ribosomal subunits with
2DTM {[}\protect\hyperlink{ref-T1Qr0xyP}{13}{]}. Due to the larger size of
mammalian cells compared to yeast cells, it was unclear whether plunge
freezing would be adequate to produce vitreous ice across the whole
volume of the cell. To test this we prepared cryo-lamellae of mouse
neutrophil cells. A low magnification image of a representative lamella
clearly shows cellular features consistent with a neutrophile-like
phenotype, mainly a segmented nucleus and a plethora of
membrane-organelles, corresponding to the granules and secretory
vesicles of neutrophils (Fig. {[}\ref{fig:initmatching}{]}A). We then proceeded
to acquire micrographs on this lamella with a defocus of 0.5-1.0 μm, 30
e\textsuperscript{-}/Å\(^2\)/s exposure and 1.76 Å pixel size. We manually selected
multiple locations in the lamella and acquired micrographs using
standard low-dose techniques where focusing is performed on a
sacrificial area. The resulting micrographs showed smooth bilayered
membranes and no signs of crystalline ice (Fig. {[}\ref{fig:initmatching}{]}C,D),
indicating successful vitrification throughout the lamella.

We used an atomic model of the 60S mouse ribosomal subunit (6SWA) for
2DTM {[}\protect\hyperlink{ref-sbkam02a}{14}{]}. In a subset of images, the
distribution of cross-correlation scores significantly exceeded the
distribution expected from images devoid of detectable targets. In the
resulting scaled maximum-intensity projections (MIPs), clear peaks with
SNR values up to 10 were apparent (Fig. {[}\ref{fig:initmatching2}{]}A). Using a
threshold criterion to select significant targets (see Methods), we
found that in images of cytosolic compartments there were 10-500
ribosomes within one micrograph (Fig. {[}\ref{fig:initmatching}{]}B-E). Notably,
we found no targets in areas corresponding to the nucleus (Fig.
{[}\ref{fig:initmatching}{]}B) or mitochondria (Fig. \protect\hyperlink{fig:initmatching}{1}D). In
the cytoplasm, we found a highly variable number of targets, only \textasciitilde{} 50
in some exposures (Fig. {[}\ref{fig:initmatching}{]}E) and up to 500 in others
(Fig. {[}\ref{fig:initmatching}{]}C). However, it is unclear whether this
ten-fold difference in local ribosome concentration is due to technical
variation, such as sample thickness, or biological variation. To
differentiate between the two we reasoned it was important to not
manually choose imaging regions and to collect larger amounts of data.
We therefore set out to collect cryo-EM data for 2DTM from mammalian
cell lamellae in a high-throughput unbiased fashion.

\hypertarget{deco-lace-for-2d-imaging-of-whole-lamellae}{%
\subsubsection{DeCo-LACE for 2D imaging of whole lamellae}\label{deco-lace-for-2d-imaging-of-whole-lamellae}}

In order to obtain high-resolution data from complete lamellae, we
developed a new approach for data collection. This approach uses three
key strategies: (1) every electron that exposes a fresh area of the
sample is collected on the camera (2) image shift is used to precisely
and quickly raster the surface of a lamella and (3) focusing is done
without using a sacrificial area (Fig. {[}\ref{fig:approach}{]}A).

To ensure that every electron exposing a fresh area of the sample is
captured by the detector, we adjusted the electron beam size to be
entirely contained by the detector area. During canonical low-dose
imaging, the microscope is configured so that the focal plane is
identical to the eucentric plane of the specimen stage. This leaves the
C2 aperture out of focus, resulting in ripples at the edge of the beam
(Fig. {[}\ref{fig:approach}{]}D). While these ripples are low-resolution features
that likely do not interfere with 2DTM {[}\protect\hyperlink{ref-Ynb3IP6I}{7}{]}, we
also tested data collection under conditions where the C2 aperture is in
focus (``fringe-free'', Fig. {[}\ref{fig:approach}{]}E)
{[}\protect\hyperlink{ref-wsFP5c54}{15}{]}.

We then centered a lamella on the optical axis of the microscope and
used the image shift controls of the microscope to systematically scan
the whole surface of the lamella in a hexagonal pattern (Fig.
{[}\ref{fig:approach}{]}A,C). Instead of focusing on a sacrificial area, we
determined the defocus from every exposure after it was taken. The
defocus was then adjusted based on the difference between desired and
measured defocus (Fig. {[}\ref{fig:approach}{]}B). Since we used a serpentine
pattern for data collection, every exposure was close to the previous
exposure, making large changes in the defocus unlikely. Furthermore, we
started our acquisition pattern on the platinum deposition edge to make
sure that the initial exposure where the defocus was not yet adjusted
did not contain any biologically relevant information.

We used this strategy to collect data on eight lamellae, four using the
eucentric focus condition, hereafter referred to as Lamella\textsubscript{EUC}, and
four using the fringe-free condition, hereafter referred to as
Lamella\textsubscript{FFF}(Fig. {[}\ref{fig:assembly}{]} A+D, Fig. {[}\ref{fig:lamella_images}{]}A). We
were able to collect data with a highly consistent defocus of 800 nm
(Fig. {[}\ref{fig:approach}{]}F), both in the eucentric focus and fringe-free
focus condition. To ensure that data were collected consistently, we
mapped defocus values as a function of the applied image shift (Fig.
{[}\ref{fig:lamella_spatial_info}{]}A). This demonstrated that the defocus was
consistent across a lamella, except for rare outliers and in images
containing contamination. We also plotted the measured objective
astigmatism of each lamella and found that it varies with the applied
image shift, becoming more astigmatic mostly due to image shift in the x
direction (Fig. {[}\ref{fig:lamella_spatial_info}{]}B). While approaches exist to
correct for this during the data collection
{[}\protect\hyperlink{ref-APSL9LmU}{16}{]}, we opted to not use these approaches
in our initial experiments. We reasoned that because 2DTM depends on
high-resolution information, this would be an excellent test of how much
these aberration affect imaging.

We assembled the tile micrographs into a montage using the image-shift
values and the SerialEM calibration followed by cross-correlation based
refinement (see Methods). In the resulting montages, the same cellular
features visible in the overview images are apparent (Fig.
{[}\ref{fig:assembly}{]}B+E, Fig. {[}\ref{fig:lamella_images}{]}B), however due to the
high magnification and low defocus many more details, such as the
membrane bilayer separation, can be observed (Fig. {[}\ref{fig:assembly}{]}C+F).
For montages collected using the eucentric condition, there are clearly
visible fringes at the edges between the tiles (Fig. {[}\ref{fig:assembly}{]}C),
which are absent in the fringe-free focus montages (Fig.
{[}\ref{fig:assembly}{]}F). In our analysis below, we show that these fringes do
not impede target detection by 2DTM, making them primarily an aesthetic
issue. We also note that the tiling pattern is visible in the montages
(Fig. {[}\ref{fig:assembly}{]}B+E), which we believe is due to the non-linear
behavior of the K3 camera since we can observe these shading artifacts
in micrographs of a condensed beam over vacuum (Fig. {[}\ref{fig:gain}{]}).

The montages show membrane vesicles and granules with highly variable
sizes and density. We found that a substantial number of granules, which
are characterized by higher density inside the the surrounding cytosol
{[}\protect\hyperlink{ref-xE2EhAcH}{17}{]}, seemed to contain a membrane-enclosed
inclusion with density similar to the surrounding cytosol (Fig.
{[}\ref{fig:lamella_images}{]}C) and could therefore be formed by inward budding
of the granule membrane. These granules were 150-300 nm in diameter and
the inclusions were 100-200 nm in diameter. Based on these dimensions
the granules are either azurophil or specific granules
{[}\protect\hyperlink{ref-xE2EhAcH}{17}{]}. To our knowledge, these inclusion have not
been described in granulocytes and are further described and discussed
below.

\hypertarget{dtm-of-deco-lace-data-reveals-large-ribosomal-subunit-distribution-in-cellular-cross-sections}{%
\subsubsection{2DTM of DeCo-LACE data reveals large ribosomal subunit distribution in cellular cross-sections}\label{dtm-of-deco-lace-data-reveals-large-ribosomal-subunit-distribution-in-cellular-cross-sections}}

In our initial attempts of using 2DTM on micrographs acquired with the
DeCo-LACE protocol, we did not observe any SNR peaks above threshold
using the large subunit of the mouse ribosome (Fig.
{[}\ref{fig:crop_unblur}{]}A). We reasoned that the edges of the beam might
interfere with motion-correction of the movies as they represent strong
low-resolution features that do not move with the sample. When we
cropped the movie frames to exclude the beam edges, the estimated amount
of motion increased (Fig. {[}\ref{fig:crop_unblur}{]}B), consistent with
successful tracking of sample motion. Furthermore, in the
motion-corrected average we could identify significant SNR peaks (Fig.
{[}\ref{fig:crop_unblur}{]}B), confirming the high sensitivity of 2DTM to the
presence of high-resolution signal preserved in the images by the motion
correction. To streamline data processing, we implemented a function in
unblur to consider only a defined central area of a movie for estimation
of sample motion, while still averaging the complete movie frames (Fig.
{[}\ref{fig:crop_unblur}{]}C). Using this approach, we motion-corrected all tiles
in the eight lamellae and found consistently total motion below 1 Å per
frame (Fig. {[}\ref{fig:lamella_motion_thickness}{]} A). In some lamellae we
found increased motion in the lamella center, which indicates areas of
variable mechanical stability within FIB-milled lamellae. In some
micrographs we also observed that the beam edges gave rise to artifacts
in the MIP and numerous false-positive detections at the edge of the
illuminated area (Fig. {[}\ref{fig:crop_unblur}{]}D). A similar phenomenon was
observed on isolated ``hot'' pixels in unilluminated areas. To overcome
this issue we implemented a function in unblur to replace dark areas in
the micrograph with Gaussian noise (see Methods), with mean and standard
deviation matching the illuminated portion of the micrograph (Fig.
{[}\ref{fig:crop_unblur}{]}D+E). Together, these pre-processing steps enabled us
to perform 2DTM on all tiles of the eight lamellae.

We used the tile positions to calculate the positions of the detected
LSUs in the lamellae (Fig. {[}\ref{fig:matching_euc}{]}A, Fig.
{[}\ref{fig:matching_fff}{]}A, Fig. {[}\ref{fig:matching_euc}{]}-video 1, Fig.
{[}\ref{fig:matching_fff}{]}-video 1). Overlaying these positions of
the lamellae montages reveals LSU distribution throughout the FIB-milled
slices of individual cells. Consistent with prior observations imaging
selected views in yeast {[}\protect\hyperlink{ref-T1Qr0xyP}{13}{]}, organelles
like the nucleus and mitochondria only showed sporadic targets detected
with low SNRs, consistent with the estimated false-positive rate of one
per tile. For each detected target we also calculated the Z positions
from the individual estimated defocus and defocus offset for each tile.
When viewed from the side, the ribosome positions therefore show the
slight tilts of the lamellae relative to the microscope frame of
reference (Fig. {[}\ref{fig:matching_euc}{]}B, Fig.
{[}\ref{fig:matching_fff}{]}B, 
Fig. {[}\ref{fig:matching_euc}{]}-video 1, Fig.
{[}\ref{fig:matching_fff}{]}-video 1). Furthermore, the side views indicated that lamellae were
thinner at the leading edge. To confirm this we estimated the ice
thickness in individual tiles using the Beer-Lambert law
{[}\protect\hyperlink{ref-17IP4Lhz1}{18}{]} with 322 nm as coefficient
{[}\protect\hyperlink{ref-17IP4Lhz1}{18}{]} (Fig.
{[}\ref{fig:lamella_motion_thickness}{]}B). We found that the thickness of the
lamellae varied between 100 and 250 nm, with consistently thinner ice at
the leading edge even though we prepared the lamellae with the
``overtilt'' approach {[}\protect\hyperlink{ref-hr3dyUeI}{19}{]}. To further confirm
the relationship of the range in Z-positions of detections and sample
thickness we plotted the range in Z-positions as a function of the
estimated sample thickness (Fig. {[}\ref{fig:lamella_motion_thickness}{]}C). For
most tiles, the apparent thickness of LSU detections was at least 70 nm
less than the estimated sample thickness, indicating that no detections
were being made within 35 nm of the lamella edge. It is possible that
this is caused by sample damage during FIB-milling as similar values for
damaged areas in FIB-milled lamellae have been estimated from
sub-tomogram averaging {[}\protect\hyperlink{ref-SDMGf7aG}{20}{]}. This means that
LSU detections across the lamellae can be skewed by a change in
thickness and sample damage. Therefore, better sample preparation
methods together with a better understanding of the damage to the sample
during FIB-milling are needed to accurately quantify molecular
concentrations.

As described in {[}\protect\hyperlink{ref-Ynb3IP6I}{7}{]} the 2DTM SNR threshold for
detecting a target is chosen to result in one false positive detection
per image searched. We would therefore expect to find one false positive
detection per tile. We reasoned that the large nuclear area imaged by
DeCo-LACE could be used to test whether this assumption is true. In the
670 tiles containing exclusively nucleus (as manually annotated from the
overview image) we detected 247 targets, making the false-positive rate
more than twofold lower than expected. Since earlier work shows that
2DTM with the LSU can produce matches to nuclear ribosome biogenesis
intermediates {[}\protect\hyperlink{ref-T1Qr0xyP}{13}{]}, this could even be an
overestimate of the false-positive rate. This suggests that the
detection threshold could be even lower, which is an area of ongoing
research.

Close inspection of the LSU positions in the lamellae revealed several
interesting features. LSUs could be seen associating with membranes, in
patterns reminiscent of the rough endoplasmic reticulum (Fig.
{[}\ref{fig:matching_euc}{]}C, Fig. {[}\ref{fig:matching_fff}{]}C) or the outer nuclear
membrane (Fig. {[}\ref{fig:matching_euc}{]}D). We also observed LSUs forming
ring-like structures (Fig. {[}\ref{fig:matching_euc}{]}E), potentially indicating
circularized mRNAs {[}\protect\hyperlink{ref-cie13Q8F}{21}{]}. While ribosomes were for the most
part excluded from the numerous granules observed in the cytoplasm, in
some cases we observed clusters of LSUs in the inclusions of
double-membraned granules described earlier (Fig. {[}\ref{fig:matching_euc}{]}F,
Fig. {[}\ref{fig:matching_fff}{]}D,E). It is, in principle, possible that these
targets are situated above or below the imaged granules, since the
granule positions in z cannot be determined using 2D projections.
However, in the case of Fig. {[}\ref{fig:matching_fff}{]}E, the detected LSUs
span the whole lamella in the z direction (Fig. {[}\ref{fig:matching_fff}{]}F),
while positions above or below a granule would result in LSUs situated
exclusively at the top or bottom of the lamella. This is consistent with
the earlier hypothesis that the inclusions are of cytoplasmic origin.

\hypertarget{does-deco-lace-induce-aberrations-that-affect-2dtm}{%
\subsubsection{Does DeCo-LACE induce aberrations that affect 2DTM?}\label{does-deco-lace-induce-aberrations-that-affect-2dtm}}

Within the eight lamellae we found different numbers of detected
targets, ranging from 1089 to 6433 per lamella (Fig.
{[}\ref{fig:matching_stat}{]}A). Lamella\textsubscript{EUC} 1 had the most detected targets,
but also has the largest surface area and contained cytoplasm from two
cells. Lamella\textsubscript{FFF} 4 had the fewest detected targets, but this
particular lamella was dominated by a circular section of the nucleus,
with only small pockets of cytoplasm (Fig. {[}\ref{fig:lamella_images}{]}). In an
attempt to normalize for these differences in area containing cytoplasm,
we compared the number of detected targets per tile in tiles that
contained more than one target, which should exclude tiles with
non-cytosolic content (Fig. {[}\ref{fig:matching_stat}{]}B). While this measure
had less variability, there were still differences. Lamella\textsubscript{EUC} 4 had
not only the fewest targets, but also the lowest density, which could be
due to this lamella being the thinnest, or due to it sectioning the cell
in an area with a lower concentration of ribosomes. Lamella\textsubscript{FFF} 3 had a
substantially higher number of ribosomes per tile. Since all of these
lamellae were made from a cell-line under identical conditions, this
underscores the necessity to collect data from large numbers of lamellae
to overcome the inherent variability. When comparing the distribution of
scores between lamellae, we found them to be fairly comparable with
median SNRs ranging from 8.7 to 9.7 (Fig. {[}\ref{fig:matching_stat}{]}C).
Lamella\textsubscript{EUC} 1 had slightly lower scores compared to the rest,
potentially due to its large size and connected mechanical instability
during imaging. Overall, we did not observe differences in the number or
SNR of detected targets between eucentric or fringe-free illumination
conditions that were bigger than the observed inter-lamella variability.

Since the SNR values of 2DTM are highly sensitive to image quality, we
reasoned we could use them to verify that DeCo-LACE does not introduce a
systematic loss of image quality. We considered non-parallel
illumination introduced by the unusually condensed beam and
uncharacterized aberrations near the beam periphery. When plotting the
SNR values of detected targets in all eight lamellae as a function of
their location in the tiles, we found uniformly high SNR values
throughout the illuminated areas for both eucentric and fringe-free
focus illumination, demonstrating that both illumination schemes are
suitable for DeCo-LACE (Fig. {[}\ref{fig:matching_stat}{]}D).

We also wondered whether large image shifts would lead to aberrations
due to astigmatism or beam tilt {[}\protect\hyperlink{ref-APSL9LmU}{16}{]}. We
reasoned that if that was the case the number of detected targets should
be highest in the center of the lamella where the applied beam
image-shift is 0. Instead, we observed that in both eucentric and
fringe-free focus conditions more targets were detected at the ``back''
edge of the lamella (Fig. {[}\ref{fig:matching_stat}{]}E{]}). This may be due to
the center of the cell being predominantly occupied by the nucleus,
despite its segmentation in neutrophil-like cells. The increase in
matches at the ``back'' of the lamellae compared to the ``front'' can also
be explained by the thickness gradient of the lamellae (Fig.
{[}\ref{fig:lamella_motion_thickness}{]}B, Fig. {[}\ref{fig:matching_euc}{]}B, Fig.
{[}\ref{fig:matching_fff}{]}B). In addition, aberrations would be expected to
cause average 2DTM SNRs to be higher when beam-image shift values are
small. Instead, we found that SNRs where on average the highest at the
``front'' edge of the lamellae, presumably due to the thinner sample. We
therefore conclude that factors other that beam image-shift or beam
condensation aberrations are limiting 2DTM SNRS, predominantly the
thickness of the lamellae.

\hypertarget{computation-is-the-bottleneck-of-visual-proteomics}{%
\subsubsection{Computation is the bottleneck of visual proteomics}\label{computation-is-the-bottleneck-of-visual-proteomics}}

All lamellae described above were derived from a clonal cell line under
identical condition and thinned with the same parameters. This means
that the substantial variability of detected targets between the
lamellae must be due to technical variability, including area,
thickness, mechanical stability, and location of the section within the
cell. We therefore predict that further studies that want to draw
quantitative and statistically relevant conclusions about the number and
location of molecules under different experimental conditions, will
require collection of orders of magnitude more data than in this study
to gain enough statistical power given this variability. The samples
used were prepared in two 24 h sessions on a FIB/SEM instrument, and
imaging was performed during another two 24h session on the TEM
microscope. Inspections of the timestamps of the raw data files revealed
that the milling time per lamella was \textasciitilde30 minutes and TEM imaging was
accomplished in \textasciitilde10 seconds per tile or 90 minutes for a \textasciitilde{} 6x6 μm
lamella. Processing of the data, however, took substantially longer.
Specifically, 2DTM of all tiles took approximately one week per lamella
on 32 Nvidia A6000 GPUs. Computation is therefore a bottleneck in our
current workflow, and further optimizations of the algorithm may be
necessary increase throughput. Alternatively, this bottleneck could be
reduced by increasing the number of processing units.

\hypertarget{discussion}{%
\subsection{Discussion}\label{discussion}}

In this study we developed an approach to image entire cellular
cross-section using cryo-EM at high enough resolution to allow for 2DTM
detection of the LSU. The two main advantages compared to previous
approaches are high throughput of imaging and biological context for
detected molecules. The requirement to increase throughput in cryo-EM
data collection of cellular samples has been recognized in the recent
literature. Most approaches described so far are tailored towards
tomography. Peck et al. {[}\protect\hyperlink{ref-1GPPSX111}{22}{]} and Yang et al.
{[}\protect\hyperlink{ref-ujizcpVn}{23}{]} developed approaches to increase the
FOV of tomogram data-collection by using a montaging technique. Peck et
al.~used a similar ``condensed-beam'' approach as described here. However,
the montages are substantially smaller in scope, covering carbon film
holes of 2 µm diameter. Bouvette et al.
{[}\protect\hyperlink{ref-1Gwx5ScUa}{24}{]} and Eisenstein et al.
{[}\protect\hyperlink{ref-xvouNMBy}{25}{]} are using beam image-shift to collect
tilt-series in multiple locations in parallel to increase throughput.
However, none of these approaches provide the full coverage of a
cellular cross-section that can be achieved using DeCo-LACE.

We observed granules containing a vesicle of putative cytosolic origin.
We speculate that upon degranulation, the process in which granules fuse
with the plasma membrane, these vesicles would be released into the
extracelullar space. The main types of extracellular vesicles of this
size are exosomes, up to 100 nm large vesicles derived from fusion of
multivesicular bodies with the plasma membrane, and microvesicles, which
are derived from direct budding of the plasma membrane
{[}\protect\hyperlink{ref-kncnYGLD}{26}{]}. We suggest that granulocytes could release
a third type of extracellular vesicle, granule-derived vesicles (GDV),
into the extracellular space. 2DTM showed that a subset of GDVs can
contain ribosomes (Fig. {[}\ref{fig:matching_euc}{]}F, Fig.
{[}\ref{fig:matching_fff}{]}D,E). This could indicate that these vesicles are
transporting translation-capable mRNAs, as has been described for
exosomes {[}\protect\hyperlink{ref-agVdpX76}{27}{]}. Further studies will be necessary to
confirm the existence of GDVs in granulocytes isolated from mammals and
to understand their functional significance.

As mentioned in the results, we found a consistent shading artifact
pattern in our montages, that we believe is the result of non-linear
behavior of the K3 camera. Indeed, when we average images with a
condensed beam taken over vacuum we found in both focus conditions a
consistent background pattern with a brighter region on the periphery of
the illuminated area (Fig {[}\ref{fig:gain}{]}). This might be caused by dynamic
adjustment of the internal camera counting threshold which expects
columns of the sensor to be evenly illuminated as is the case for SPA
applications. Since the signal of this pattern has mainly low-resolution
components it is unlikely to affect 2DTM. However, it highlights that
the non-linear behavior of the camera has to be taken into account when
imaging samples with strongly varying density and unusual illumination
schemes.

We found that even though we used beam image-shift extensively (up to 7
um), we did not see substantially reduced 2DTM SNR values in tiles
acquired at high beam image-shift compared to tiles acquired with low or
no beam image-shift. This is unexpected because beam image-shift is
though to degrade the high-resolution signal due to the introduction of
beam-tilt, as has been observed in single particle analysis
{[}\protect\hyperlink{ref-44fG6kcy}{28},\protect\hyperlink{ref-wy8cRzrH}{29}{]}. The
signal detected by 2DTM is target and sample dependent and has been
shown to be sensitive to a resolution up to 3 Å, but in practice is
limited by the B-factor applied to the template
{[}\protect\hyperlink{ref-Ynb3IP6I}{7}{]}. In our case, we used a structure of the
mouse LSU that was deposited into the PDB with atoms having a median
B-factor of \textasciitilde{} 80 Å\(^2\), which we multiplied by 1.5 before generating
the density map. This value was chosen empirically early in the study to
optimize the number of detections. The resulting median B-factor of 120
Å\(^2\) is equivalent to a Gaussian lowpass filter with a filter constant
of 7.7 Å. We did not measure the beam-tilt caused by beam-image shift on
the microscope to acquire our data, but Cheng et. al.~report that in
their measurements Krios Titan columns exhibited a beam-tilt of
0.17-0.25 mrad per μm of beam image-shift
{[}\protect\hyperlink{ref-wy8cRzrH}{29}{]}. We therefore expect a worst-case
beam-tilt of 1.75 mrad. At this beam-tilt we expect a phase error of
larger than π/4 at resolutions higher than 5.75 Å
{[}\protect\hyperlink{ref-wy8cRzrH}{29}{]}, which exceeds the resolution of our
2DTM template. We conclude that the beam-tilt caused by beam-image shift
did not strongly affect detections in the present study, but it might be
important to consider when imaging at even higher beam-image shift
values or when using higher resolution signal in 2DTM.

We used 2DTM of the ribosome LSU to verify the image quality of
DeCo-LACE data. However, 2DTM has the potential to detect many other
protein complexes. The model of the ribosomal LSU used in this
manuscript has a molecular weight of \textasciitilde2 MDa, \textasciitilde1.3 MDa of which is
nucleic acid. The higher scattering density of nucleic acid compared to
protein {[}\protect\hyperlink{ref-13fhsbtxo}{30}{]} may mean that a biomolecular complex of equal
size consisting of only protein would result in lower SNR values. Many
biological complexes are smaller and do not contain nucleic acid, but
simulations of 2DTM {[}\protect\hyperlink{ref-Ynb3IP6I}{7}{]} suggest that it should be
possible to detect protein molecules as small as 300 kDa in an crowded
environment. Besides the molecular weight we expect that the
conformational homogeneity of a protein complex will affect the ability
to detect it and note that for complexes that occur at low
concentrations inside the cell a higher detection threshold might be
necessary to reduce the number of false positives among the few true
targets that can be detected.

Since we observed substantial variation in LSU density within and
between lamellae, visual proteomics studies that use cryo-EM to
establish changes in molecular organization within cells will require
orders of magnitude more data than used in this study. One milestone
would be to image enough data to represent one cellular volume, which
for a small eukaryotic cells requires imaging approximately 100
lamellae. While data collection throughput on the TEM is fundamentally
limited by the exposure time, this amount of data could be collected
within 12 hours by improving the data acquisition scheme to perform all
necessary calculations in parallel with actual exposure of the camera.
Sample preparation using a FIB/SEM is also currently a bottleneck, but
preparation of large lamellae with multiple cellular cross-sections
using methods like WAFFLE {[}\protect\hyperlink{ref-17H2LWiIu}{31}{]} might allow
sufficient throughput. As stated in the results, at least for 2DTM
computation will remain challenging and approximately 17,000 GPU hours
would be required for a 100 lamellae dataset. It might be possible to
increase the performance of 2DTM by improving the implementation of
GPU-based fast Fourier transforms. Alternatively, strategies that
initially search images at lower resolution and sparser rotational
sampling, followed by refinement at higher resolution, could be more
efficient, but it is yet unclear whether this is feasible in cellular
samples with strong low-resolution background.

\hypertarget{materials-and-methods}{%
\subsection{Materials and Methods}\label{materials-and-methods}}

\hypertarget{key-resources-table}{%
\subsubsection{Key Resources Table}\label{key-resources-table}}

\begin{tablenos:no-prefix-table-caption}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2139}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1148}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2451}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1148}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2951}}@{}}
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Reagent type
(species)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Designation
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Source or reference
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Identifiers
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Additional Information
\end{minipage} \\
\midrule()
\endhead
cell line (\emph{mus
musculus}) & ER-HoxB8 & {[}\protect\hyperlink{ref-1B9Vt9eYu}{10}{]} & & Generated from mouse bone marrow as
described in
{[}\protect\hyperlink{ref-1B9Vt9eYu}{10}{]} \\
peptide, recombinant protein & SCF & {[}\protect\hyperlink{ref-1B9Vt9eYu}{10}{]} & & Generated from CHO-SCF cell line as
described in
{[}\protect\hyperlink{ref-1B9Vt9eYu}{10}{]} \\
chemical compound,
drug & Estrogen & Sigma & E8875 & \\
\bottomrule()
\end{longtable}

\end{tablenos:no-prefix-table-caption}

\hypertarget{grid-preparation}{%
\subsubsection{Grid preparation}\label{grid-preparation}}

ER-HoxB8 cells were generated as described by Wang et. al.
{[}\protect\hyperlink{ref-1B9Vt9eYu}{10},\protect\hyperlink{ref-nSMAxqTD}{32}{]}
and maintained in RPMI medium supplemented with 10\% FBS,
penicillin/streptomycin, SCF, and estrogen
 at 37 °C and 5\% CO2. 120 h prior to grid preparation, cells were washed twice in PBS and
cultured in the same medium except without estrogen. The authenticity of the
cells was verified by flow-cytometry and Giemsa staining after differentiation {[}\protect\hyperlink{ref-1B9Vt9eYu}{10},\protect\hyperlink{ref-nSMAxqTD}{32}{]}. Cells were then counted and diluted to \(1\cdot10^6\)
cells/ml. Grids (either 200 mesh copper grids, with a sillicone-oxide
and 2 µm holes with a 2 µm spacing or 200 mesh gold grids with a thin
gold film and 2 µm holes in 2 µm spacing) were glow-discharged from both
sides using a 15 mA for 45 s. 3.5 µl of cell suspension was added to
grids on the thin-film side and grids were blotted from the back side
using a GP2 cryoplunger (Leica, Wetzlar, Germany) for 8 s and rapidly
plunged into liquid ethane at -185 °C.

\hypertarget{fib-milling}{%
\subsubsection{FIB-milling}\label{fib-milling}}

Grids were loaded into an Aquilos 2 FIB/SEM (Thermo Fisher, Waltham, MA)
instrument with a stage cooled to -190 °C. Grids were sputter-coated
with platinum for 15 s at 45 mA and then coated with a layer of
platinum-precursor by opening the GIS-valve for 45 s. An overview of the
grid was created by montaging SEM images and isolated cells at the
center of gridsquares were selected for FIB-milling. Lamellae were
generated automatically using the AutoTEM software (Thermo Fisher), with
the following parameters:

\begin{itemize}
\tightlist
\item
  Milling angle: 20°
\item
  Rough milling: 3.2 µm thickness, 0.5 nA current
\item
  Medium milling: 1.8 µm thickness, 0.3 nA current, 1.0° overtilt
\item
  Fine milling: 1.0 µm tchickness, 0.1 nA current, 0.5° overtilt
\item
  Finer milling: 700 nm thickness, 0.1 nA curent, 0.2° overtilt
\item
  Polish 1: 450 nm thickness, 50 pA current
\item
  Polish 2: 200 nm thickness, 30 pA current
\end{itemize}

This resulted in 6-10 µm wide lamella with 150-250 nm thickness as
determined by FIB-imaging of the lamella edges.

\hypertarget{data-collection}{%
\subsubsection{Data collection}\label{data-collection}}

Grids were loaded into a Titan Krios TEM (Thermo Fisher) operated at 300
keV and equipped with a BioQuantum energy filter (Gatan, Pleasanton, CA)
and K3 camera (Gatan). The microscope was aligned using a cross-grating
grid on the stage. Prior to each session, we carefully performed the
``Image/Beam'' calibration in nanoprobe. We set the magnification to a
pixel size of 1.76 Å and condensed the beam to \textasciitilde{} 900 nm diameter,
resulting in the beam being completely visible on the camera. To
establish fringe-free conditions, the ``Fine eucentric'' procedure of
SerialEM {[}\protect\hyperlink{ref-19ZFerhph}{33}{]} was used to move a square of
the cross-grating grid to the eucentric position of the microscope. The
effective defocus was then set to 2 µm, using the ``autofocus'' routine of
SerialEM. The objective focus of the microscope was changed until no
fringes were visible. The stage was then moved in Z until images had an
apparent defocus of 2 µm. The difference in stage Z-position between the
eucentric and fringe-free conditions was used to move other areas into
fringe-free condition.

Low magnification montages were used to find lamellae and lamellae that
were sufficiently thin and free of contamination were selected for
automated data collection. Overview images of each lamella were taken at
2250x magnification (38 Å pixel size). The corners of the lamella in the
overview image were manually annotated in SerialEM and translated into
beam image-shift values using SerialEM's calibration. A hexagonal
pattern of beam image-shift positions was calculated that covered the
area between the four corners in a serpentine way, with a
\(\sqrt{3}\cdot 0.95 \cdot r\) horizontal spacing and
\(3/2\cdot 0.95 \cdot r\) vertical spacing, where \(r\) is the radius of the
illuminated area. For the fringe-free focus condition r was 250 nm and
for the eucentric focus r was 230 nm. This resulted in a fractional
overlap of adjacent tiles of 10.3\% for the fringe-free focus condition
and 12.9 \% for the eucentric focus condition. Exposures were taken at
each position with a 30 e\textsuperscript{-}/Å\(^2\) total dose. After each exposure, the
defocus was estimated using the ctffind function of SerialEM and the
focus for the next exposure was corrected by the difference between the
estimated focus and the desired defocus of 800 nm. Furthermore, after
each exposure the deviation of the beam from the center of the camera
was measured and corrected using the ``CenterBeamFromImage'' command of
SerialEM.

After data collection, a 20 s exposure at 2250x magnification of the
lamella at 200 μm defocus was taken for visualization purposes. A Python
script implementing this procedure is available at
\url{https://github.com/jojoelfe/deco_lace_template_matching_manuscript}.

\hypertarget{deco-lace-data-processing}{%
\subsubsection{DeCo-LACE data processing}\label{deco-lace-data-processing}}

An overview of the data analysis pipeline is shown in Fig.
\ref{fig:deco_lace_workflow}.

\hypertarget{pre-processing}{%
\paragraph{Pre-processing}\label{pre-processing}}

Motion-correction, dose weighting and other preprocessing as detailed
below was performed using \emph{cis}TEM {[}\protect\hyperlink{ref-MUY42yxp}{34}{]}. To avoid
influence of the beam-edge on motion-correction, only a quarter of the
movie in the center of the camera was considered for calculation of the
estimated motion. After movie frames were aligned and summed, a mask for
the illuminated area was calculated by lowpass filtering the image with
a 100 Å resolution cutoff, thresholding the image at 10\% of the maximal
value and then lowpass filtering the mask again with a 100 Å resolution
cutoff to smooth the mask edges. This mask was then used to select dark
areas in the image and fill the pixels with Gaussian noise, with the
same mean and standard deviation as the illuminated area. A custom
version of the unblur program {[}\protect\hyperlink{ref-1G4Y94qCy}{35}{]} implementing
this procedure is available at {[}link to decolace branch{]}. During
motion correction images were resampled to a pixel size of 1.5 Å. The
contrast-transfer function (CTF) was estimated using ctffind
{[}\protect\hyperlink{ref-n43f3Sqs}{36}{]}, searching between 0.2 and 2 μm
defocus.

\hypertarget{dtm}{%
\paragraph{2DTM}\label{dtm}}

The search template was generated from the atomic model of the mouse LSU
(PDB 6SWA, exluding the Epb1 subunit) using the cryo-EM simulator
implemented in \emph{cis}TEM {[}\protect\hyperlink{ref-q2PSamkS}{37}{]}. The
match\_template program {[}\protect\hyperlink{ref-10bXZuF3G}{9}{]} was used to search for
this template in the movie-aligned, exposure-filtered and masked images,
using a 1.5° angular step in out-of-plane angles and a 1.0° angular step
in-plane. 11 defocus planes in 20 nm steps centered around the
ctffind-determined defocus were searched. Targets were defined as
detected when their matches with the template produced peaks with a
singal-to-noise ratio (SNR) above a threshold of 7.75, which was chosen
based on the one-false-positive-per-tile criterion
{[}\protect\hyperlink{ref-Ynb3IP6I}{7}{]}.

\hypertarget{montage-assembly}{%
\paragraph{Montage assembly}\label{montage-assembly}}

The coordinates of each tile \(i\), \(\mathbf{c}_{i}\) {[}2D Vector in
pixels{]} were initialized using beam image-shift of the tile,
\(\mathbf{b}_i\) {[}2D Vector in μm{]}, and the ISToCamera matrix
\(\mathbf{IC}\), as calibrated by SerialEM:

\[\mathbf{c}_{i} = \mathbf{IC} \cdot \mathbf{b}_i\]

A list of tile pairs \(i,j\) that overlap were assembled by selecting
images where \(|\mathbf{c}_i-\mathbf{c}_j| < D_{Beam}\). In order to
calculate the precise offset between tiles \(i\) and \(j\),
\(\mathbf{r}_{i,j}\), we calculated the cross-correlation between the two
tiles, masked to the overlapping illuminated area using the scikit-image
package {[}\protect\hyperlink{ref-stvWEJeu}{38}{]} was used to calculate refined offsets .
The coordinates \(\mathbf{c}_{i}\) were then refined by a least-square
minimization against \(\mathbf{r}_{i,j}\):

\[ \displaystyle{\min_{\mathbf{c}} \sum_{pairs}{(\mathbf{r}_{i,j} - (\mathbf{c}_i-\mathbf{c}_j))^2}}\]

using the scipy package {[}\protect\hyperlink{ref-8Miti2Gz}{39}{]}. The masked
cross-correlation and the least-square minimization was repeated once
more to arrive at the final tile alignment.

The x,y coordinates of target \(n\) detected by 2DTM in the tile \(i\),
\(\textbf{m}^\textrm{T}_{n,i}\), was transformed into the montage frame by
adding the coordinate of the tile.

\[ \textbf{m}^\textrm{M}_n = \textbf{m}^\textrm{T}_{n,i} + \textbf{c}_i\]

The z coordinate of each target was calculated as the sum of the defocus
offset for the target, the estimated defocus of the tile, and the
nominal defocus of the microscope when the tile was acquired.

Images were rendered using UCSF ChimeraX {[}\protect\hyperlink{ref-cYCuuZxG}{40}{]} using a
custom extension to render 2DTM results available at
\url{https://github.com/jojoelfe/tempest}. The Python scripts used for data
processing are available under
\url{https://github.com/jojoelfe/deco_lace_template_matching_manuscript}.

\hypertarget{acknowledgments}{%
\subsection{Acknowledgments}\label{acknowledgments}}

The authors would like to thank Bronwyn Lucas, Carsten Sachse, and Chen
Xu for helpful suggestions and careful reading of the manuscriptas as
well as members of the Grigorieff lab for helpful discussions. Data was
collected at the UMass Chan Medical School Cryo-EM core with help by
Kankang Song, Christna Ouch, and Chen Xu.

\hypertarget{data-availability}{%
\subsection{Data availability}\label{data-availability}}

Cryo-EM movies, motion-corrected images and 2DTM results have been
deposited in EMPIAR under accession code EMPIAR-11063. The custom cisTEM
version is available under
\url{https://github.com/jojoelfe/cisTEM/tree/2574dbdf6161658fd177660b3a841100a792f61b}
until features have been integrated into a cisTEM release. The ChimeraX
extension for rendering is available under
\url{https://github.com/jojoelfe/tempest}. This manuscript was prepared
using the manubot package {[}\protect\hyperlink{ref-YuJbg3zO}{41}{]} . The
corresponding repository containing all scripts used for figure
generation can be found under
\url{https://github.com/jojoelfe/deco_lace_template_matching_manuscript}.

\hypertarget{conflicts-of-interest}{%
\subsection{Conflicts of interest}\label{conflicts-of-interest}}

The Authors declare that there is no conflict of interest.

\hypertarget{figures}{%
\subsection{Figures}\label{figures}}

\begin{figure}
\hypertarget{fig:initmatching}{%
\centering
\includegraphics{figures/initmatching.png}
\caption{2D template matching of the large subunit of the ribosome in
fib-milled neutrophil-like cells (A) Overview image of the lamella.
Major cellular regions are labeled, as Nucleus (Nuc), Mitochondria (M),
and granular cytoplasm (GrCyt). FOVs where high-magnification images for
template matching where acquired are indicated as boxes with the number
of detected targets indicated on the bottom right. FOVs displayed in
Panels B-E are color-coded. Scalebar corresponds to 1 μm. (B-E) FOVs
with projection of detected LSUs shown in cyan. (B) Perinuclear region,
the only detected targets are in the cytoplasmic half. (C) Cytoplasmic
region with high density of ribosomes (D) Mitochondrium, as expected
there are only detected LSUs in the cytoplasmic region (E) Cytoplasm,
with low density of
ribosomes.}\label{fig:initmatching}
}
\end{figure}

\begin{fignos:tagged-figure}[1 - figure supplement 1]

\begin{figure}
\hypertarget{fig:initmatching2}{%
\centering
\includegraphics{figures/initmatching2.png}
\caption{2D template matching of the large subunit of the ribosome in
fib-milled neutrophil-like cells (A) Maximum intensity projection (MIP)
cross-correlation map of micrograph shown in Figure \ref{fig:initmatching}
(B+C) 3D plot of MIP regions indicated by color boxes in Panel
A}\label{fig:initmatching2}
}
\end{figure}

\end{fignos:tagged-figure}

\begin{figure}
\hypertarget{fig:approach}{%
\centering
\includegraphics{figures/approach.png}
\caption{DeCo-LACE approach (A) Graphic demonstrating the data-collection
strategy for DeCo-LACE. The electron beam is condensed to a diameter
\(D_{Beam}\) that allows captured of the whole illuminated area on the
camera. Beam-image shift along X and Y (\(BIS_X\),\$\(BIS_Y\)) is used to
scan the whole lamella (B) Diagram of the collection algorithm (C)
Example overview image of a lamella with the designated acquisition
positions and the used beam diameter indicated with red circles.
Scalebar corresponds to 1 μm. (D+E) Representative micrographs takne
with a condensed beam at eucentric focus (D) or fringe-free focus (E).
Scalebar corresponds to 100 nm. (F) Boxplot of defocus measured by
ctffind of micrographs taken by the DeCo-LACE approach on four lamellae
images at eucentric focus and four lamellae imaged with fringe-free
focus.}\label{fig:approach}
}
\end{figure}

\begin{fignos:tagged-figure}[2 - figure supplement 1]

\begin{figure}
\hypertarget{fig:lamella_spatial_info}{%
\centering
\includegraphics{figures/lamella_spatial_info.png}
\caption{Defocus estimation of individual tiles of DeCo-LACE montages (A)
Defocus values of individual micrographs taken using the DeCo-LACE
approach plotted as a function of the beam image-shift values. (B)
Defocus astigmatism of individual micrographs taken using the DeCo-LACE
approach plotted as a function of the beam image-shift
values.}\label{fig:lamella_spatial_info}
}
\end{figure}

\end{fignos:tagged-figure}

\begin{figure}
\hypertarget{fig:assembly}{%
\centering
\includegraphics{figures/assembly.png}
\caption{Assembling DeCo-LACE exposures into montages (A) Overview image of
Lamella\textsubscript{EUC} 1 taken at low magnification. Scalebar corresponds to 1 μm.
(B) Overview of Lamella\textsubscript{EUC} 1 created by montaging high magnification
images taken with the DeCo-LACE approach. Scalebar corresponds to 1 μm.
(C) Zoom-in into red box in panel B. Slight beam-fringe artifacts are
visible. Scalebar corresponds to 100 nm. (D) Overview image of
Lamella\textsubscript{FFF} 4 taken at low magnification. Scalebar corresponds to 1 μm.
(E) Overview of Lamella\textsubscript{FFF} 4 created by montaging high magnification
images taken with the DeCo-LACE approach. Scalebar corresponds to 1 μm.
(F) Zoom-in into red box in panel E. No beam-fringe artifacts are
visible. Scalebar corresponds to 100
nm.}\label{fig:assembly}
}
\end{figure}

\begin{fignos:tagged-figure}[3 - figure supplement 1]

\begin{figure}
\hypertarget{fig:crop_unblur}{%
\centering
\includegraphics[width=\textwidth,height=18cm]{figures/crop_unblur.png}
\caption{Motion correction of movies with condensed beams. At the top of each
panel is an average of the movie that was motion-corrected with a red
dashed box indicating the region that was used to estimate shifts. Below
is a graph indicating the estimated shifts of the individual frames of
the movie. Below this is the MIP of 2DTM using the large subunit of the
mouse ribosome. (A) Motion correction of the whole movie (B) Notion
correction of a cropped region of the movie that eliminates the beam
edges (C) Motion correction of the whole movie, using only the central
region to estimate the shifts (D) Representative result of 2DTM (bottom)
on uncorrected micrographs. Detections along the beam edge are artifacts
due to the high-contrast signal. (E) Micrographs after preprocessing for
2DTM (top). The micrograph has been cropped to the perimeter of the beam
and unilluminated areas of the image have been replaced with Gaussian
noise. After running 2DTM on this micrograph (bottom), no more
artifactual detections are
observed.}\label{fig:crop_unblur}
}
\end{figure}

\end{fignos:tagged-figure}

\begin{fignos:tagged-figure}[3 - figure supplement 2]

\begin{figure}
\hypertarget{fig:lamella_motion_thickness}{%
\centering
\includegraphics[width=\textwidth,height=20cm]{figures/lamella_motion_thickness.png}
\caption{Motion correction of individual tiles imaged using the DeCo-LACE
approach (A) Total estimated motion of individual micrographs taken
using the DeCo-LACE approach plotted as a function of the beam
image-shift values. (B) Sample thickness estimate of individual
micrographs taken using the DeCo-LACE approach plotted as a function of
the beam image-shift values. Sample thickness was estimated using
Lamber-Beers law using a coefficient of 322 nm
{[}\protect\hyperlink{ref-17IP4Lhz1}{18}{]}. (C) Distance between highest and
lowest LSU detection in each micrograph plotted as a function of
estimated sample thickness. Dashed black line indicates unity and red
dashed line is drawn in parallel with an x intercept of 70 nm. Most data
points fall below the red dashed line, indicating that no detections
occur within 35 nm of the lamella
edge.}\label{fig:lamella_motion_thickness}
}
\end{figure}

\end{fignos:tagged-figure}

\begin{fignos:tagged-figure}[3 - figure supplement 3]

\begin{figure}
\hypertarget{fig:gain}{%
\centering
\includegraphics{figures/gain.png}
\caption{Averages of micrographs taken with a condensed beam over vacuum using
a Gatan K3 detector. Contrast and Brightness have been adjusted to
highlight uneven dose response. (A) Eucentric Focus (B) Fringe-free
Focus}\label{fig:gain}
}
\end{figure}

\end{fignos:tagged-figure}

\begin{fignos:tagged-figure}[3 - figure supplement 4]

\begin{figure}
\hypertarget{fig:lamella_images}{%
\centering
\includegraphics[width=\textwidth,height=23cm]{figures/lamella_images.png}
\caption{Overview images of lamellae imaged using the DeCo-LACE approach taken
at low-magnification (A) Overviews taken at low magnification. Scalebar
corresponds to 1 μm. (B) Overviews assembled using the DeCo-LACE
approach. Scalebar corresponds to 1 μm. (C) Representative examples of a
class of granules containing a putatively cytosolic inclusion. Scalebar
corresponds to 100 nm.}\label{fig:lamella_images}
}
\end{figure}

\end{fignos:tagged-figure}

\begin{figure}
\hypertarget{fig:matching_euc}{%
\centering
\includegraphics{figures/matching_euc.png}
\caption{Template matching in lamella imaged using the DeCo-LACE approach at
eucentric focus (A) Montage of Lamella\(_\textrm{EUC}\) 1 overlaid with
detected targets colored in orange. Scalebar corresponds to 1 μm. (B)
Side view of detected targets in the lamella, such that the direction of
the electron beam is horizontal. (C-F) Magnified area of panel A showing
rough ER with associated ribosomes (C), outer nuclear membrane with
associated ribosomes (D), ribsomes arranged in a circular fashion (E),
ribosomes enclosed in a less electron dense inclusion in a granule (F).
Ribosomes are colored in white with the surface of the peptide exit
tunnel colored in green and the A, P, and E sites colored in blue,
purple, and red, respectively.Scalebar corresponds to 100
nm.}\label{fig:matching_euc}
}
\end{figure}

\begin{figure}
\hypertarget{fig:matching_fff}{%
\centering
\includegraphics{figures/matching_fff.png}
\caption{Template matching in lamella imaged using the DeCo-LACE approach at
fringe-free focus (A) Montage of Lamella\(_\textrm{FFF}\) 4 overlaid with
detected targets colored in orange. Scalebar corresponds to 1 μm. (B)
Side view of detected targets in the lamella, such that the direction of
the electron beam is horizontal. (C-E) Magnified area of panel A showing
rough ER with associated ribosomes (C) and ribosomes enclosed in a less
electron dense inclusion in a granule (D,E). (F) Side view of panel E.
Ribosomes are colored in white with the surface of the peptide exit
tunnel colored in green and the A, P, and E sites colored in blue,
purple, and red, respectively. Scalebar corresponds to 100
nm.}\label{fig:matching_fff}
}
\end{figure}

\begin{figure}
\hypertarget{fig:matching_stat}{%
\centering
\includegraphics{figures/matching_stat.png}
\caption{Statistics of 2DTM on lamella imaged using DeCo-LACE (A) Number of
detected targets in each lamella (B) Distribution of targets per tile in
each lamella. Only tiles with two or more detected targets were included
(C) Distribution of SNRs in each lamella (D) For each lamella an average
of all tiles is shown. Overlaid is a scatterplot of all detected targets
in these tiles according to their in-tile coordinates. Scatterplot is
colored according to the 2DTM SNR. There are no detected targets in the
top circle-circle intersection due to radiation damage from previous
exposures. (E) 2D histogram of number of detected targets as a function
of beam-image shift (F) Mean 2DTM SNR as a function of beam-image
shift}\label{fig:matching_stat}
}
\end{figure}

\begin{figure}
\hypertarget{fig:deco_lace_workflow}{%
\centering
\includegraphics{figures/deco_lace_workflow.png}
\caption{Workflow of DeCo-LACE
processing}\label{fig:deco_lace_workflow}
}
\end{figure}

\begin{fignos:tagged-figure}[Figure 4-video 1]

\begin{figure}
\hypertarget{fig:movie1}{%
\centering
\includegraphics{images/1x1-00000000.png}
\caption{Movie of detected LSU targets in Lamella\(_\textrm{EUC}\) 1,
corresponding to Figure 4}\label{fig:movie1}
}
\end{figure}

\end{fignos:tagged-figure}

\begin{fignos:tagged-figure}[Figure 5-video 1]

\begin{figure}
\hypertarget{fig:movie2}{%
\centering
\includegraphics{images/1x1-00000000.png}
\caption{Movie of detected LSU targets in Lamella\(_\textrm{FFF}\) 4,
corresponding to Figure 5}\label{fig:movie2}
}
\end{figure}

\end{fignos:tagged-figure}

\hypertarget{references}{%
\subsection{References}\label{references}}

\hypertarget{refs}{}
\begin{CSLReferences}{0}{0}
\leavevmode\vadjust pre{\hypertarget{ref-tSXIKPl7}{}}%
\CSLLeftMargin{1. }%
\CSLRightInline{\textbf{Label-free, normalized quantification of complex mass spectrometry data for proteomic analysis}
\CSLBlock{Noelle M Griffin, Jingyi Yu, Fred Long, Phil Oh, Sabrina Shore, Yan Li, Jim A Koziol, Jan E Schnitzer} \emph{Nature Biotechnology} (2010-01) \url{https://doi.org/fshgnc}
\CSLBlock{DOI: \href{https://doi.org/10.1038/nbt.1592}{10.1038/nbt.1592} · PMID: \href{https://www.ncbi.nlm.nih.gov/pubmed/20010810}{20010810} · PMCID: \href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC2805705}{PMC2805705}}}

\leavevmode\vadjust pre{\hypertarget{ref-VBmW7Aot}{}}%
\CSLLeftMargin{2. }%
\CSLRightInline{\textbf{Fluorescence microscopy}
\CSLBlock{Jeff W Lichtman, José-Angel Conchello} \emph{Nature Methods} (2005-11-18) \url{https://doi.org/bbpg4n}
\CSLBlock{DOI: \href{https://doi.org/10.1038/nmeth817}{10.1038/nmeth817} · PMID: \href{https://www.ncbi.nlm.nih.gov/pubmed/16299476}{16299476}}}

\leavevmode\vadjust pre{\hypertarget{ref-tGQ6TSUo}{}}%
\CSLLeftMargin{3. }%
\CSLRightInline{\textbf{A visual approach to proteomics}
\CSLBlock{Stephan Nickell, Christine Kofler, Andrew P Leis, Wolfgang Baumeister} \emph{Nature Reviews Molecular Cell Biology} (2006-02-15) \url{https://doi.org/d6d5mq}
\CSLBlock{DOI: \href{https://doi.org/10.1038/nrm1861}{10.1038/nrm1861} · PMID: \href{https://www.ncbi.nlm.nih.gov/pubmed/16482091}{16482091}}}

\leavevmode\vadjust pre{\hypertarget{ref-g8QavfwP}{}}%
\CSLLeftMargin{4. }%
\CSLRightInline{\textbf{Electron microscopy of frozen hydrated sections of vitreous ice and vitrified biological samples}
\CSLBlock{AW McDowall, J-J Chang, R Freeman, J Lepault, CA Walter, J Dubochet} \emph{Journal of Microscopy} (1983-07) \url{https://doi.org/bdnzmv}
\CSLBlock{DOI: \href{https://doi.org/10.1111/j.1365-2818.1983.tb04225.x}{10.1111/j.1365-2818.1983.tb04225.x} · PMID: \href{https://www.ncbi.nlm.nih.gov/pubmed/6350598}{6350598}}}

\leavevmode\vadjust pre{\hypertarget{ref-16IhS1Nc4}{}}%
\CSLLeftMargin{5. }%
\CSLRightInline{\textbf{Opening windows into the cell: focused-ion-beam milling for cryo-electron tomography}
\CSLBlock{Elizabeth Villa, Miroslava Schaffer, Jürgen M Plitzko, Wolfgang Baumeister} \emph{Current Opinion in Structural Biology} (2013-10) \url{https://doi.org/f537jp}
\CSLBlock{DOI: \href{https://doi.org/10.1016/j.sbi.2013.08.006}{10.1016/j.sbi.2013.08.006} · PMID: \href{https://www.ncbi.nlm.nih.gov/pubmed/24090931}{24090931}}}

\leavevmode\vadjust pre{\hypertarget{ref-Rksh2dxu}{}}%
\CSLLeftMargin{6. }%
\CSLRightInline{\textbf{Electron tomography of cells}
\CSLBlock{Lu Gan, Grant J Jensen} \emph{Quarterly Reviews of Biophysics} (2011-11-15) \url{https://doi.org/czj9hr}
\CSLBlock{DOI: \href{https://doi.org/10.1017/s0033583511000102}{10.1017/s0033583511000102} · PMID: \href{https://www.ncbi.nlm.nih.gov/pubmed/22082691}{22082691}}}

\leavevmode\vadjust pre{\hypertarget{ref-Ynb3IP6I}{}}%
\CSLLeftMargin{7. }%
\CSLRightInline{\textbf{Single-protein detection in crowded molecular environments in cryo-EM images}
\CSLBlock{JPeter Rickgauer, Nikolaus Grigorieff, Winfried Denk} \emph{eLife} (2017-05-03) \url{https://doi.org/gnq4q4}
\CSLBlock{DOI: \href{https://doi.org/10.7554/elife.25648}{10.7554/elife.25648} · PMID: \href{https://www.ncbi.nlm.nih.gov/pubmed/28467302}{28467302} · PMCID: \href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5453696}{PMC5453696}}}

\leavevmode\vadjust pre{\hypertarget{ref-18KGpXYPE}{}}%
\CSLLeftMargin{8. }%
\CSLRightInline{\textbf{Label-free single-instance protein detection in vitrified cells}
\CSLBlock{JPeter Rickgauer, Heejun Choi, Jennifer Lippincott-Schwartz, Winfried Denk} \emph{Cold Spring Harbor Laboratory} (2020-04-24) \url{https://doi.org/gpbjfd}
\CSLBlock{DOI: \href{https://doi.org/10.1101/2020.04.22.053868}{10.1101/2020.04.22.053868}}}

\leavevmode\vadjust pre{\hypertarget{ref-10bXZuF3G}{}}%
\CSLLeftMargin{9. }%
\CSLRightInline{\textbf{Locating macromolecular assemblies in cells by 2D template matching with cisTEM}
\CSLBlock{Bronwyn A Lucas, Benjamin A Himes, Liang Xue, Timothy Grant, Julia Mahamid, Nikolaus Grigorieff} \emph{eLife} (2021-06-11) \url{https://doi.org/gkkc49}
\CSLBlock{DOI: \href{https://doi.org/10.7554/elife.68946}{10.7554/elife.68946} · PMID: \href{https://www.ncbi.nlm.nih.gov/pubmed/34114559}{34114559} · PMCID: \href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC8219381}{PMC8219381}}}

\leavevmode\vadjust pre{\hypertarget{ref-1B9Vt9eYu}{}}%
\CSLLeftMargin{10. }%
\CSLRightInline{\textbf{Inhibition of Dihydroorotate Dehydrogenase Overcomes Differentiation Blockade in Acute Myeloid Leukemia}
\CSLBlock{David B Sykes, Youmna S Kfoury, François E Mercier, Mathias J Wawer, Jason M Law, Mark K Haynes, Timothy A Lewis, Amir Schajnovitz, Esha Jain, Dongjun Lee, \ldots{} David T Scadden} \emph{Cell} (2016-09) \url{https://doi.org/f3r5jr}
\CSLBlock{DOI: \href{https://doi.org/10.1016/j.cell.2016.08.057}{10.1016/j.cell.2016.08.057} · PMID: \href{https://www.ncbi.nlm.nih.gov/pubmed/27641501}{27641501} · PMCID: \href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7360335}{PMC7360335}}}

\leavevmode\vadjust pre{\hypertarget{ref-gRoY21jY}{}}%
\CSLLeftMargin{11. }%
\CSLRightInline{\textbf{Hallmarks of ribosomopathies}
\CSLBlock{Kim R Kampen, Sergey O Sulima, Stijn Vereecke, Kim De~Keersmaecker} \emph{Nucleic Acids Research} (2019-07-27) \url{https://doi.org/gpbjfm}
\CSLBlock{DOI: \href{https://doi.org/10.1093/nar/gkz637}{10.1093/nar/gkz637} · PMID: \href{https://www.ncbi.nlm.nih.gov/pubmed/31350888}{31350888} · PMCID: \href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7026650}{PMC7026650}}}

\leavevmode\vadjust pre{\hypertarget{ref-KAJ7221k}{}}%
\CSLLeftMargin{12. }%
\CSLRightInline{\textbf{Diagnostic and prognostic implications of ribosomal protein transcript expression patterns in human cancers}
\CSLBlock{James M Dolezal, Arie P Dash, Edward V Prochownik} \emph{BMC Cancer} (2018-03-12) \url{https://doi.org/gc87j9}
\CSLBlock{DOI: \href{https://doi.org/10.1186/s12885-018-4178-z}{10.1186/s12885-018-4178-z} · PMID: \href{https://www.ncbi.nlm.nih.gov/pubmed/29530001}{29530001} · PMCID: \href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5848553}{PMC5848553}}}

\leavevmode\vadjust pre{\hypertarget{ref-T1Qr0xyP}{}}%
\CSLLeftMargin{13. }%
\CSLRightInline{\textbf{ In situ single particle classification reveals distinct 60S maturation intermediates in cells}
\CSLBlock{Bronwyn A Lucas, Kexin Zhang, Sarah Loerch, Nikolaus Grigorieff} \emph{Cold Spring Harbor Laboratory} (2022-04-10) \url{https://doi.org/gp4qrx}
\CSLBlock{DOI: \href{https://doi.org/10.1101/2022.04.10.487797}{10.1101/2022.04.10.487797}}}

\leavevmode\vadjust pre{\hypertarget{ref-sbkam02a}{}}%
\CSLLeftMargin{14. }%
\CSLRightInline{\textbf{Protein Synthesis in the Developing Neocortex at Near-Atomic Resolution Reveals Ebp1-Mediated Neuronal Proteostasis at the 60S Tunnel Exit}
\CSLBlock{Matthew L Kraushar, Ferdinand Krupp, Dermot Harnett, Paul Turko, Mateusz C Ambrozkiewicz, Thiemo Sprink, Koshi Imami, Manuel Günnigmann, Ulrike Zinnall, Carlos H Vieira-Vieira, \ldots{} Christian MT Spahn} \emph{Molecular Cell} (2021-01) \url{https://doi.org/gh7d72}
\CSLBlock{DOI: \href{https://doi.org/10.1016/j.molcel.2020.11.037}{10.1016/j.molcel.2020.11.037} · PMID: \href{https://www.ncbi.nlm.nih.gov/pubmed/33357414}{33357414} · PMCID: \href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC8163098}{PMC8163098}}}

\leavevmode\vadjust pre{\hypertarget{ref-wsFP5c54}{}}%
\CSLLeftMargin{15. }%
\CSLRightInline{\textbf{Advances in Single Particle Analysis Data Acquisition}
\CSLBlock{Stan Konings, Maarten Kuijper, Jeroen Keizer, Fanis Grollios, Tjerk Spanjer, Peter Tiemeijer} \emph{Microscopy and Microanalysis} (2019-08) \url{https://doi.org/gp9bqm}
\CSLBlock{DOI: \href{https://doi.org/10.1017/s1431927619005798}{10.1017/s1431927619005798}}}

\leavevmode\vadjust pre{\hypertarget{ref-APSL9LmU}{}}%
\CSLLeftMargin{16. }%
\CSLRightInline{\textbf{High-quality, high-throughput cryo-electron microscopy data collection via beam tilt and astigmatism-free beam-image shift}
\CSLBlock{Chunling Wu, Xiaojun Huang, Jing Cheng, Dongjie Zhu, Xinzheng Zhang} \emph{Journal of Structural Biology} (2019-12) \url{https://doi.org/gp2nq9}
\CSLBlock{DOI: \href{https://doi.org/10.1016/j.jsb.2019.09.013}{10.1016/j.jsb.2019.09.013} · PMID: \href{https://www.ncbi.nlm.nih.gov/pubmed/31562921}{31562921}}}

\leavevmode\vadjust pre{\hypertarget{ref-xE2EhAcH}{}}%
\CSLLeftMargin{17. }%
\CSLRightInline{\textbf{THE DEVELOPMENT OF NEUTROPHILIC POLYMORPHONUCLEAR LEUKOCYTES IN HUMAN BONE MARROW}
\CSLBlock{Dorothy Ford Bainton, Joan L Ullyot, Marilyn G Farquhar} \emph{Journal of Experimental Medicine} (1971-10-01) \url{https://doi.org/chkbcm}
\CSLBlock{DOI: \href{https://doi.org/10.1084/jem.134.4.907}{10.1084/jem.134.4.907} · PMID: \href{https://www.ncbi.nlm.nih.gov/pubmed/4106490}{4106490} · PMCID: \href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC2138991}{PMC2138991}}}

\leavevmode\vadjust pre{\hypertarget{ref-17IP4Lhz1}{}}%
\CSLLeftMargin{18. }%
\CSLRightInline{\textbf{Routine determination of ice thickness for cryo-EM grids}
\CSLBlock{William J Rice, Anchi Cheng, Alex J Noble, Edward T Eng, Laura Y Kim, Bridget Carragher, Clinton S Potter} \emph{Journal of Structural Biology} (2018-10) \url{https://doi.org/gfbpj2}
\CSLBlock{DOI: \href{https://doi.org/10.1016/j.jsb.2018.06.007}{10.1016/j.jsb.2018.06.007} · PMID: \href{https://www.ncbi.nlm.nih.gov/pubmed/29981485}{29981485} · PMCID: \href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6119488}{PMC6119488}}}

\leavevmode\vadjust pre{\hypertarget{ref-hr3dyUeI}{}}%
\CSLLeftMargin{19. }%
\CSLRightInline{\textbf{Optimized cryo-focused ion beam sample preparation aimed at in situ structural studies of membrane proteins}
\CSLBlock{Miroslava Schaffer, Julia Mahamid, Benjamin D Engel, Tim Laugks, Wolfgang Baumeister, Jürgen M Plitzko} \emph{Journal of Structural Biology} (2017-02) \url{https://doi.org/f9qfr6}
\CSLBlock{DOI: \href{https://doi.org/10.1016/j.jsb.2016.07.010}{10.1016/j.jsb.2016.07.010} · PMID: \href{https://www.ncbi.nlm.nih.gov/pubmed/27444390}{27444390}}}

\leavevmode\vadjust pre{\hypertarget{ref-SDMGf7aG}{}}%
\CSLLeftMargin{20. }%
\CSLRightInline{\textbf{Plasma FIB milling for the determination of structures \textless i\textgreater in situ\textless/i\textgreater{}}
\CSLBlock{Casper Berger, Maud Dumoux, Thomas Glen, Neville B-y Yee, John M Mitchels, Zuzana Patáková, James H Naismith, Michael Grange} \emph{Cold Spring Harbor Laboratory} (2022-08-01) \url{https://doi.org/gq59b8}
\CSLBlock{DOI: \href{https://doi.org/10.1101/2022.08.01.502333}{10.1101/2022.08.01.502333}}}

\leavevmode\vadjust pre{\hypertarget{ref-cie13Q8F}{}}%
\CSLLeftMargin{21. }%
\CSLRightInline{\textbf{Circularization of mRNA by eukaryotic translation initiation factors.}
\CSLBlock{SE Wells, PE Hillner, RD Vale, AB Sachs} \emph{Molecular cell} (1998-07) \url{https://www.ncbi.nlm.nih.gov/pubmed/9702200}
\CSLBlock{DOI: \href{https://doi.org/10.1016/s1097-2765(00)80122-7}{10.1016/s1097-2765(00)80122-7} · PMID: \href{https://www.ncbi.nlm.nih.gov/pubmed/9702200}{9702200}}}

\leavevmode\vadjust pre{\hypertarget{ref-1GPPSX111}{}}%
\CSLLeftMargin{22. }%
\CSLRightInline{\textbf{Montage electron tomography of vitrified specimens}
\CSLBlock{Ariana Peck, Stephen D Carter, Huanghao Mai, Songye Chen, Alister Burt, Grant J Jensen} \emph{Journal of Structural Biology} (2022-06) \url{https://doi.org/gp4qrs}
\CSLBlock{DOI: \href{https://doi.org/10.1016/j.jsb.2022.107860}{10.1016/j.jsb.2022.107860} · PMID: \href{https://www.ncbi.nlm.nih.gov/pubmed/35487464}{35487464}}}

\leavevmode\vadjust pre{\hypertarget{ref-ujizcpVn}{}}%
\CSLLeftMargin{23. }%
\CSLRightInline{\textbf{Correlative cryogenic montage electron tomography for comprehensive in-situ whole-cell structural studies}
\CSLBlock{Jie E Yang, Matthew R Larson, Bryan S Sibert, Joseph Y Kim, Daniel Parrell, Juan C Sanchez, Victoria Pappas, Anil Kumar, Kai Cai, Keith Thompson, Elizabeth R Wright} \emph{Cold Spring Harbor Laboratory} (2022-01-02) \url{https://doi.org/gp4qrv}
\CSLBlock{DOI: \href{https://doi.org/10.1101/2021.12.31.474669}{10.1101/2021.12.31.474669}}}

\leavevmode\vadjust pre{\hypertarget{ref-1Gwx5ScUa}{}}%
\CSLLeftMargin{24. }%
\CSLRightInline{\textbf{Beam image-shift accelerated data acquisition for near-atomic resolution single-particle cryo-electron tomography}
\CSLBlock{Jonathan Bouvette, Hsuan-Fu Liu, Xiaochen Du, Ye Zhou, Andrew P Sikkema, Juliana da Fonseca Rezende e Mello, Bradley P Klemm, Rick Huang, Roel M Schaaper, Mario J Borgnia, Alberto Bartesaghi} \emph{Nature Communications} (2021-03-30) \url{https://doi.org/gjphkb}
\CSLBlock{DOI: \href{https://doi.org/10.1038/s41467-021-22251-8}{10.1038/s41467-021-22251-8} · PMID: \href{https://www.ncbi.nlm.nih.gov/pubmed/33785757}{33785757} · PMCID: \href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC8009872}{PMC8009872}}}

\leavevmode\vadjust pre{\hypertarget{ref-xvouNMBy}{}}%
\CSLLeftMargin{25. }%
\CSLRightInline{\textbf{Parallel cryo electron tomography on \textless i\textgreater in situ\textless/i\textgreater{} lamellae}
\CSLBlock{Fabian Eisenstein, Haruaki Yanagisawa, Hiroka Kashihara, Masahide Kikkawa, Sachiko Tsukita, Radostin Danev} \emph{Cold Spring Harbor Laboratory} (2022-04-08) \url{https://doi.org/gp4qrw}
\CSLBlock{DOI: \href{https://doi.org/10.1101/2022.04.07.487557}{10.1101/2022.04.07.487557}}}

\leavevmode\vadjust pre{\hypertarget{ref-kncnYGLD}{}}%
\CSLLeftMargin{26. }%
\CSLRightInline{\textbf{Shedding light on the cell biology of extracellular vesicles}
\CSLBlock{Guillaume van Niel, Gisela D'Angelo, Graça Raposo} \emph{Nature Reviews Molecular Cell Biology} (2018-01-17) \url{https://doi.org/gct4df}
\CSLBlock{DOI: \href{https://doi.org/10.1038/nrm.2017.125}{10.1038/nrm.2017.125} · PMID: \href{https://www.ncbi.nlm.nih.gov/pubmed/29339798}{29339798}}}

\leavevmode\vadjust pre{\hypertarget{ref-agVdpX76}{}}%
\CSLLeftMargin{27. }%
\CSLRightInline{\textbf{Exosome-mediated transfer of mRNAs and microRNAs is a novel mechanism of genetic exchange between cells}
\CSLBlock{Hadi Valadi, Karin Ekström, Apostolos Bossios, Margareta Sjöstrand, James J Lee, Jan O Lötvall} \emph{Nature Cell Biology} (2007-05-07) \url{https://doi.org/d5df4s}
\CSLBlock{DOI: \href{https://doi.org/10.1038/ncb1596}{10.1038/ncb1596} · PMID: \href{https://www.ncbi.nlm.nih.gov/pubmed/17486113}{17486113}}}

\leavevmode\vadjust pre{\hypertarget{ref-44fG6kcy}{}}%
\CSLLeftMargin{28. }%
\CSLRightInline{\textbf{High-resolution cryo-EM using beam-image shift at 200 keV}
\CSLBlock{Jennifer N Cash, Sarah Kearns, Yilai Li, Michael A Cianfrocco} \emph{IUCrJ} (2020-10-29) \url{https://doi.org/gjwcfq}
\CSLBlock{DOI: \href{https://doi.org/10.1107/s2052252520013482}{10.1107/s2052252520013482} · PMID: \href{https://www.ncbi.nlm.nih.gov/pubmed/33209328}{33209328} · PMCID: \href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7642776}{PMC7642776}}}

\leavevmode\vadjust pre{\hypertarget{ref-wy8cRzrH}{}}%
\CSLLeftMargin{29. }%
\CSLRightInline{\textbf{High resolution single particle cryo-electron microscopy using beam-image shift}
\CSLBlock{Anchi Cheng, Edward T Eng, Lambertus Alink, William J Rice, Kelsey D Jordan, Laura Y Kim, Clinton S Potter, Bridget Carragher} \emph{Journal of Structural Biology} (2018-11) \url{https://doi.org/gfg6vz}
\CSLBlock{DOI: \href{https://doi.org/10.1016/j.jsb.2018.07.015}{10.1016/j.jsb.2018.07.015} · PMID: \href{https://www.ncbi.nlm.nih.gov/pubmed/30055234}{30055234} · PMCID: \href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6163078}{PMC6163078}}}

\leavevmode\vadjust pre{\hypertarget{ref-13fhsbtxo}{}}%
\CSLLeftMargin{30. }%
\CSLRightInline{\textbf{A method for differentiating proteins from nucleic acids in intermediate-resolution density maps: cryo-electron microscopy defines the quaternary structure of the Escherichia coli 70S ribosome.}
\CSLBlock{CM Spahn, PA Penczek, A Leith, J Frank} \emph{Structure (London, England : 1993)} (2000-09-15) \url{https://www.ncbi.nlm.nih.gov/pubmed/10986461}
\CSLBlock{DOI: \href{https://doi.org/10.1016/s0969-2126(00)00185-4}{10.1016/s0969-2126(00)00185-4} · PMID: \href{https://www.ncbi.nlm.nih.gov/pubmed/10986461}{10986461}}}

\leavevmode\vadjust pre{\hypertarget{ref-17H2LWiIu}{}}%
\CSLLeftMargin{31. }%
\CSLRightInline{\textbf{Waffle Method: A general and flexible approach for improving throughput in FIB-milling}
\CSLBlock{Kotaro Kelley, Ashleigh M Raczkowski, Oleg Klykov, Pattana Jaroenlak, Daija Bobe, Mykhailo Kopylov, Edward T Eng, Gira Bhabha, Clinton S Potter, Bridget Carragher, Alex J Noble} \emph{Nature Communications} (2022-04-06) \url{https://doi.org/gp4qrt}
\CSLBlock{DOI: \href{https://doi.org/10.1038/s41467-022-29501-3}{10.1038/s41467-022-29501-3} · PMID: \href{https://www.ncbi.nlm.nih.gov/pubmed/35387991}{35387991} · PMCID: \href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC8987090}{PMC8987090}}}

\leavevmode\vadjust pre{\hypertarget{ref-nSMAxqTD}{}}%
\CSLLeftMargin{32. }%
\CSLRightInline{\textbf{Quantitative production of macrophages or neutrophils ex vivo using conditional Hoxb8}
\CSLBlock{Gang G Wang, Katherine R Calvo, Martina P Pasillas, David B Sykes, Hans Häcker, Mark P Kamps} \emph{Nature Methods} (2006-03-22) \url{https://doi.org/dgxn4m}
\CSLBlock{DOI: \href{https://doi.org/10.1038/nmeth865}{10.1038/nmeth865} · PMID: \href{https://www.ncbi.nlm.nih.gov/pubmed/16554834}{16554834}}}

\leavevmode\vadjust pre{\hypertarget{ref-19ZFerhph}{}}%
\CSLLeftMargin{33. }%
\CSLRightInline{\textbf{Automated electron microscope tomography using robust prediction of specimen movements}
\CSLBlock{David N Mastronarde} \emph{Journal of Structural Biology} (2005-10) \url{https://doi.org/ff7gzx}
\CSLBlock{DOI: \href{https://doi.org/10.1016/j.jsb.2005.07.007}{10.1016/j.jsb.2005.07.007} · PMID: \href{https://www.ncbi.nlm.nih.gov/pubmed/16182563}{16182563}}}

\leavevmode\vadjust pre{\hypertarget{ref-MUY42yxp}{}}%
\CSLLeftMargin{34. }%
\CSLRightInline{\textbf{cisTEM, user-friendly software for single-particle image processing}
\CSLBlock{Timothy Grant, Alexis Rohou, Nikolaus Grigorieff} \emph{eLife} (2018-03-07) \url{https://doi.org/gf5cq7}
\CSLBlock{DOI: \href{https://doi.org/10.7554/elife.35383}{10.7554/elife.35383} · PMID: \href{https://www.ncbi.nlm.nih.gov/pubmed/29513216}{29513216} · PMCID: \href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5854467}{PMC5854467}}}

\leavevmode\vadjust pre{\hypertarget{ref-1G4Y94qCy}{}}%
\CSLLeftMargin{35. }%
\CSLRightInline{\textbf{Measuring the optimal exposure for single particle cryo-EM using a 2.6 Å reconstruction of rotavirus VP6}
\CSLBlock{Timothy Grant, Nikolaus Grigorieff} \emph{eLife} (2015-05-29) \url{https://doi.org/gf5cm7}
\CSLBlock{DOI: \href{https://doi.org/10.7554/elife.06980}{10.7554/elife.06980} · PMID: \href{https://www.ncbi.nlm.nih.gov/pubmed/26023829}{26023829} · PMCID: \href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4471936}{PMC4471936}}}

\leavevmode\vadjust pre{\hypertarget{ref-n43f3Sqs}{}}%
\CSLLeftMargin{36. }%
\CSLRightInline{\textbf{CTFFIND4: Fast and accurate defocus estimation from electron micrographs}
\CSLBlock{Alexis Rohou, Nikolaus Grigorieff} \emph{Journal of Structural Biology} (2015-11) \url{https://doi.org/f7xzgv}
\CSLBlock{DOI: \href{https://doi.org/10.1016/j.jsb.2015.08.008}{10.1016/j.jsb.2015.08.008} · PMID: \href{https://www.ncbi.nlm.nih.gov/pubmed/26278980}{26278980} · PMCID: \href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6760662}{PMC6760662}}}

\leavevmode\vadjust pre{\hypertarget{ref-q2PSamkS}{}}%
\CSLLeftMargin{37. }%
\CSLRightInline{\textbf{Cryo-TEM simulations of amorphous radiation-sensitive samples using multislice wave propagation}
\CSLBlock{Benjamin Himes, Nikolaus Grigorieff} \emph{IUCrJ} (2021-09-30) \url{https://doi.org/gpzs9r}
\CSLBlock{DOI: \href{https://doi.org/10.1107/s2052252521008538}{10.1107/s2052252521008538} · PMID: \href{https://www.ncbi.nlm.nih.gov/pubmed/34804546}{34804546} · PMCID: \href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC8562658}{PMC8562658}}}

\leavevmode\vadjust pre{\hypertarget{ref-stvWEJeu}{}}%
\CSLLeftMargin{38. }%
\CSLRightInline{\textbf{scikit-image: image processing in Python}
\CSLBlock{Stéfan van der Walt, Johannes L Schönberger, Juan Nunez-Iglesias, François Boulogne, Joshua D Warner, Neil Yager, Emmanuelle Gouillart, Tony Yu} \emph{PeerJ} (2014-06-19) \url{https://doi.org/gftp3s}
\CSLBlock{DOI: \href{https://doi.org/10.7717/peerj.453}{10.7717/peerj.453} · PMID: \href{https://www.ncbi.nlm.nih.gov/pubmed/25024921}{25024921} · PMCID: \href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4081273}{PMC4081273}}}

\leavevmode\vadjust pre{\hypertarget{ref-8Miti2Gz}{}}%
\CSLLeftMargin{39. }%
\CSLRightInline{\textbf{SciPy 1.0: fundamental algorithms for scientific computing in Python}
\CSLBlock{Pauli Virtanen, Ralf Gommers, Travis E Oliphant, Matt Haberland, Tyler Reddy, David Cournapeau, Evgeni Burovski, Pearu Peterson, Warren Weckesser, Jonathan Bright, \ldots{} } \emph{Nature Methods} (2020-02-03) \url{https://doi.org/ggj45f}
\CSLBlock{DOI: \href{https://doi.org/10.1038/s41592-019-0686-2}{10.1038/s41592-019-0686-2} · PMID: \href{https://www.ncbi.nlm.nih.gov/pubmed/32015543}{32015543} · PMCID: \href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7056644}{PMC7056644}}}

\leavevmode\vadjust pre{\hypertarget{ref-cYCuuZxG}{}}%
\CSLLeftMargin{40. }%
\CSLRightInline{\textbf{UCSF ChimeraX
: Structure visualization for researchers, educators, and developers}
\CSLBlock{Eric F Pettersen, Thomas D Goddard, Conrad C Huang, Elaine C Meng, Gregory S Couch, Tristan I Croll, John H Morris, Thomas E Ferrin} \emph{Protein Science} (2020-10-22) \url{https://doi.org/ghr6mn}
\CSLBlock{DOI: \href{https://doi.org/10.1002/pro.3943}{10.1002/pro.3943} · PMID: \href{https://www.ncbi.nlm.nih.gov/pubmed/32881101}{32881101} · PMCID: \href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7737788}{PMC7737788}}}

\leavevmode\vadjust pre{\hypertarget{ref-YuJbg3zO}{}}%
\CSLLeftMargin{41. }%
\CSLRightInline{\textbf{Open collaborative writing with Manubot}
\CSLBlock{Daniel S Himmelstein, Vincent Rubinetti, David R Slochower, Dongbo Hu, Venkat S Malladi, Casey S Greene, Anthony Gitter} \emph{PLOS Computational Biology} (2019-06-24) \url{https://doi.org/c7np}
\CSLBlock{DOI: \href{https://doi.org/10.1371/journal.pcbi.1007128}{10.1371/journal.pcbi.1007128} · PMID: \href{https://www.ncbi.nlm.nih.gov/pubmed/31233491}{31233491} · PMCID: \href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6611653}{PMC6611653}}}

\end{CSLReferences}

\end{document}